\externalReferent{core}
\section*{Rules Summary}

\begin{multicols}{2}

\subsection{Character Creation}

\begin{itemize}

  \item
  Roll a race (page \pageref{raceRoll}).
  \item
  Roll your 6 Attributes.
  \item
  Pick a concept (2 words).
  \item
  Pick a Code (page \pageref{gods_codes})
  \item
  Spend 50 XP.
  \item
  Pick 1 item per Skill (worth 10sp or less)

\end{itemize}

\subsection{Actions}

\subsubsection{Standard Actions -- page \pageref{basicaction}}

For most actions:

\begin{itemize}

  \item
  The player says ``I jump over the wall''.
  \item
  The GM replies ``Roll 2D6 and add your Speed + Athletics'' at Target Number 9.
  \item
  If the roll gets 9 or more, the action succeeds.

\end{itemize}

\subsubsection{Group Rolls -- page \pageref{grouproll}}

When many characters do the same thing:

\begin{itemize}

  \item
  One player rolls 2D6.
  \item
  Everyone adds their own Attribute + Skill to the same roll to get their own result.
  \item
  Some characters will succeed, while others fail.

\end{itemize}

\subsubsection{Team Actions -- page \pageref{teamwork}}

When working together helps:

\begin{itemize}

  \item
  One player rolls, as above.
  \item
  The character with the highest score adds to the roll.
  \item
  The second highest score adds half their score, the third adds a quarter, and so on.
  \item
  Everyone wins or loses.

\end{itemize}

\subsubsection{Resisted Actions -- page \pageref{resistedactions}}

When enemies push back:

\begin{itemize}

  \item{The basic TN is usually 7.}
  \item{The opponent roll adds their Attribute + Skill to the TN.}

\end{itemize}

\subsubsection{Resting Actions -- page \pageref{restingactions}}

If you can take your time doing something:

\begin{itemize}

  \item{Set one of your dice to `6' and roll the other one.}
  
\end{itemize}

\subsection{Combat -- page \pageref{attack}}

\begin{itemize}
  \item
  Attack by rolling 2D6 plus your Dexterity + Combat + Weapon Bonus.
  \item
  Roll your Damage dice at the same time: 1D6 + Strength + Weapon Bonus (pro tip: make your Damage dice a different colour).
  \item
  If you have more than 3 Damage, replace 4 Damage with a D6.
  \item
  If you miss, the enemy hits you!
  Mark off some FP to avoid Damage.
  \item
  If you have no FP left, mark off HP.
  \item
  Spend the AP for your weapon (if you go below 0 AP, you will take a penalty).
  \item
  Spend 1 AP to move or speak.
\end{itemize}

\subsection{Magic -- page \pageref{basiccasting}}

\begin{itemize}

  \item
  Spend MP equal to the spell's level.
  \item
  Roll Intelligence + some Skill.
  \item
  Normal spells take 1 round per level.
  \item
  Fast spells are cast at +1 level, and take 1 AP + Level - Wits.

\end{itemize}

\subsection{Stories -- page \pageref{stories}}

Spend your 5 Story Points on introducing elements from your characters history.

\begin{itemize}

  \item
  Declare you know a language.
  \item
  Introduce a friend with specialist knowledge (no Combat Skill).
  \item
  Introduce an old ally.
  \item
  When someone dies, they take one of those allies as their new character.

\end{itemize}

\subsection{Scene Ends}

\begin{itemize}

  \item
  Regain Fate Points equal to 40\% of your base.
  \item
  \textbf{Spellcasters:} Regain 2 mana points + Wits Bonus.
  \item
  \textbf{When resting:} Heal Fatigue Points equal to half your current HP.
  \item
  \textbf{End of Session:} Heal HP, and spend XP.

\end{itemize}

\end{multicols}
