\chapter{Established Ways}
\label{coreRules}

\settoggle{genExamples}{true}

\section{Resolution}
\label{basicaction}

\begin{multicols}{2}

\newcommand{\TNChart}{

  \noindent
  \begin{boxtable}[cX]

    \textbf{\glsentrytext{tn}} & \textbf{Task} \\\hline

    2 & Automatic \\

    4 & Trivial \\

    6 & Easy \\

    8 & Serious \\

    10 & Tricky \\

    12 & Professional \\

    14 & Specialist \\

    16 & Extreme \\

    18 & Legendary \\

    20 & Implausible \\

  \end{boxtable}
}

\subsection{Basic Actions}
When \glspl{pc} attempt something dangerous and difficult, the \gls{gm} states the \gls{tn}, and the players try to beat it by rolling $2D6$ plus any Bonuses.
The more difficult the action, the higher the \gls{tn}.

Players add one of their character's Attributes and Skills to the roll, and sometimes a Bonus for equipment.

\begin{multicols}{2}
\begin{itemize}

  \item
  If the player rolls above the \gls{tn}, they succeed!
  \item
  If they roll below, they fail, and the danger occurs.
  \item
  If they roll equal to the \gls{tn}, the \gls{gm} can give them the choice to succeed, as long as they accept the danger.

\end{itemize}

\columnbreak

\noindent
\TNChart

\end{multicols}

\begin{exampletext}
  Hugi listens carefully at the keyhole, trying to figure out what the elves on the other side are plotting.
  The \gls{tn} is 10, and he matches it exactly.

  ``One of the elves sounds like he's wandering closer to the door, but you think you almost heard a familiar word'', the \gls{gm} says.

  ``I'll stay and listen'', Hugi's player says.

  ``They use your name, though you can't understand the elvish -- just your name -- then one opens the door, saying `well here he sits, or stands, I can never tell with dwarves'.
  The other elves stand up quickly.''

\end{exampletext}

Assume all actions have \pgls{tn} of 7 unless your \gls{gm} states otherwise.
Don't ask -- just roll!%
\footnote{\Glspl{gm} never forget to state the \gls{tn}, so if you're told to roll, but don't hear any \gls{tn}, then the \gls{tn} \emph{must} be 7.
This is legally binding.}

These standard rules should cover any situation, with the right interpretation.
However, for suggested interpretations with more detail, see \nameref{skill_uses}, \vpageref{skill_uses}.

\subsection{One Roll Only}
\index{Group actions}

Players only make one \gls{natural} per action.
If a player wants to re-try an action, the result remains the same unless circumstances change.

When many characters are trying to do the same task, one player rolls, and all players consult the same results.

\begin{exampletext}
\begin{itemize}
  \item
  Everyone wants to kick in the door, the \gls{tn} is 10, and the roll uses \roll{Strength}{Crafts}.
  \index{Doors!Teamwork}
    \begin{itemize}
    \item
    Pikerudd's player rolls \twoDice{7}.
    Her \roll{Strength}{Crafts} Bonus is 0, so she fails.
    \item
    Snowblight's \roll{Strength}{Crafts} total is +2, so his total is 9, and he also fails.
    \item
    Chatrik's \roll{Strength}{Crafts} total is +4, so her total is 11 -- she succeeds.
    \end{itemize}
  \item
  These look like the famous Catacombs from the national anthem, but how do the words go?
  Did the hero take two lefts and a right, or two rights and a left?
    \begin{itemize}
    \item
    Water is filling the area, so time will be lost if they get muddled.
    \item
    Everyone rolls, and the highest result is Dzo's -- but he only equals the \gls{tn}.
    \item
    Dzo's player decides to guess left, rather let his companions think about the song for another moment.
    They fail, but avoid wasting time.
    \end{itemize}
\end{itemize}

\end{exampletext}

If the troupe are all attempting the same action, then they only make one roll, while adding different \gls{attribute} to obtain their individual result.

\subsection{\Glsfmtplural{restingaction}}
\label{restingactions}

Difficult, but safe actions allow players to repeat the same task until they get it right or give up.
In these cases, the player sets one die to a `6', and rolls only the other die.

\begin{exampletext}
\begin{itemize}
  \item
    The troupe really need a favour from a local \gls{warden}, but he has no time to meet with \gls{guard}~\glspl{fodder}.
  \begin{itemize}
    \item
    The \gls{gm} disallows this as \pgls{restingaction}.
    \item
    After the first failure, the noble tells his servants not to pay them any more attention -- they can no longer succeed via the official channels.
  \end{itemize}
  \item
  The group want to sneak into a noble's house, and have plenty of time to plan the heist.
  The \gls{gm} says they can wander past without suspicion, but it will take a week to plan and gather all the information they need.
  The roll is \roll{Intelligence}{Stealth} at \tn[12].
    \begin{itemize}
    \item
    The players accept, and roll: \dicef{4}.
    \item
    With \dicef{6}~\dicef{4} on the table, the dice show `10', and the total is `13' -- a narrow success.
    \end{itemize}
\end{itemize}
\end{exampletext}

\subsection{Resisted Actions}
\label{resistedactions}

When \pgls{pc} and \pgls{npc} act in opposition to each other, the player adds their characters \gls{attribute} + \gls{skill} Bonuses as usual, and the \gls{tn} equals \tn[7] plus the \gls{npc}'s \gls{attribute} + \gls{skill}.%
\footnote{In general, only players roll for actions.}

\begin{exampletext}
  \begin{itemize}
    \item
      A local cutthroat wants to sneak up on \pgls{pc} while they go shopping for armour, and corner them alone.
      Her \roll{Wits}{Stealth} gives her a +3 Bonus, so the \gls{gm} asks for a \roll{Wits}{Vigilance} roll at \tn[10].
    \item
      The troupe need to convince the locals not to trust a particular psychotic, lying, bastard.
      Despite him not being present, the \gls{gm} notes that he his \roll{Charisma}{Empathy} grants a +5 Bonus, so he states the \gls{tn} is 12.
    \item
    \Pgls{witch} begins a curse against a member of the \gls{guard}, so he decides to stab her.
    Her \roll{Charisma}{Fate} total is +2, so the \gls{tn} is ($7+2$) 9.

    The guard uses his \roll{Dexterity}{Melee} to resist the curse, so if he wins, he will inflict Damage (covered in \autoref{combat}).
    But if the \gls{witch}'s spell works, the \gls{guard} will fumble his attack, and lose some \glspl{fp}.
  \end{itemize}
\end{exampletext}

\subsection{\Glsfmtplural{bandAct}}
\label{teamwork}
\label{banding}
\index{Teamwork}

Some tasks lend themselves to working with others.
Others can be difficult or impossible to do with companions.
When the troupe want to work together to get a broken cart down a hill safely, track down a thief, or spot danger, they can benefit from \pgls{bandAct}.
But when sneaking through an area, navigating, or understanding history, Banding together will not help.

When characters can work together, one person rolls with their Bonus, then adds half the second Bonus, a quarter of the next Bonus, and so on.%
\footnote{Always round up on half, so $4 + \frac{1}{2} = 5$, but $5 + \frac{1}{4} = 5$.}

\begin{exampletext}
  Convincing the townsfolk that they need to rebel against \pgls{warden}, the troupe work together on a \roll{Charisma}{Combat} roll.
  Pikerudd has the highest Bonus, so he adds the full Bonus, then Snowblight adds half of his, and Chatrik, a quarter.

  \begin{boxtable}[L|ccc]
                            & Pikerudd & Snowblight      & Chatrik         \\
  \hline                                                                   
  Charisma + Combat         &  +4     &     +2           & +2              \\
  Multiplier                &   1     & $\frac{1}{2}$    & $\frac{1}{4}$   \\
  Value:                    &   4     & $1$              & $\frac{1}{2}$   \\
  Running Total:            &   4     & $5$              & $5+\frac{1}{2}$ \\
  \hline
    Grand Total: & & & \textbf{6} \\
  \end{boxtable}
\end{exampletext}

\index{Searching}
Working together doesn't have to mean staying together.
When searching \pgls{area}, \pgls{bandAct} generally involves splitting up.

\subsubsection{Generalized \Glsfmtplural{bandAct}}
means all accumulated Bonuses work the same way.
If two people are pushing with Strength +2, they count as having a total Strength of +3.
You'll find a few different rules are just the \gls{bandAct} rule applied to a different area (for example, \nameref{bandingArmour}, \vpageref{bandingArmour}).

\subsection{Margins}
\index{Margins}
\index{Failure Margin}
\label{margin}

If you ever need detail on how well an action went, look at how many points above the \gls{tn} the dice show.
With \pgls{tn} of 12, rolling 14 means a Margin of~2.

The \gls{gm} might use a Margin for some variable, for example a bard attempting to charm a crowd into giving him money might gain $2D6$ copper pieces plus the Margin, so if the Margin is 3 then he would get $2D6+3$ copper pieces.
Margins might also be used to gain Bonuses on later rolls.
Someone attempting to impress a noble court might roll \roll{Charisma}{Combat}; the bigger the Margin the more troops they will be trusted with, granting a Bonus to later strategy rolls.

`Failure Margins' measure how far someone falls short of \pgls{tn}.
When the \gls{tn} to attack someone is `12', then rolling `9' means a Failure Margin of `3', which often means that \gls{armour} counts for nothing, and the opponent scores \pgls{vitalShot}.

\subsection{What the Dice Mean}

You might think of the dice as representing random chance in the environment. Just how irritated is that person you're trying to question, and how creative is that craftsman feeling today? Dice are never re-rolled for different results on the same action because once the dice have told you what the situation is, the situation stays put.

Such a do-over still suggests initial failure; it just means that the character is trying over and over again until they obtain a better result.
Characters cannot roll dice for a different result unless the situation has changed notably, and generally not during the same day.

\end{multicols}

\section{\Glsfmtname{weight}}
\label{weight}
\index{Encumbrance}

\begin{multicols}{2}

\noindent%
\Glspl{weight} track how much characters can carry in broad terms.
Small items, like a dagger or \gls{ration} have \pgls{weight} of~1,
a small tent or longsword could have \pgls{weight} of 2, and bulky items, like a cartwheel have a higher \gls{weight}, to reflect difficulty carrying them.

Once a character carries a total \gls{weight} equal to their current \glspl{hp}, they take a -1~Penalty to all \glspl{attribute}.

\begin{exampletext}
  Sind\"e puts on her chainmail armour, and she's already regretting joining the \gls{guard}.
  It's heavy, and occupies her first two `\glsfmttext{hp}-slots'.
  The shortsword has \pgls{weight} of 1, so that occupies the third place, then a few coins in the fourth.
  At this point she can move fine, without Penalty; but once she picks up \pgls{ration} for the day, the total \gls{weight} equals her \glspl{hp}, and she takes a -1~Penalty to all \glspl{attribute}.
  She won't be fighting so well or \gls{running} so fast.

\begin{boxtable}[YYYYY|c]
           1  &              2  &      3  &       4  &     5   & Penalty  \\\hline
  chain-- & --mail  &  \footnotesize shortsword  & 50~\glspl{cp}  &  \footnotesize\rations & \textbf{-1} \\
  %\rations &  \glsentrysymbol{weight} &  \glsentrysymbol{weight}  &  \glsentrysymbol{weight} & \glsentrysymbol{weight} & \textbf{-2} \\
\end{boxtable}

\end{exampletext}

\begin{exampletext}
  Battling \pgls{basilisk} leaves Sind\"e with a broken and bleeding arm.
  She only has 2~\glspl{hp} left, and carries \pgls{weight} of 4 in total, so she's carrying \pgls{weight} equal to her \glspl{hp} twice, which means a -2~Penalty to all \glspl{attribute}.
\end{exampletext}

\begin{boxtable}[YYYYY|c]
             1  &              2  &     3     &     4   &   5    & Penalty  \\\hline
    chain-- & --mail              &     X     &     X   &   X    &           \\
    \footnotesize shortsword  & 50~\glspl{cp}  &           &     &    & \textbf{-2} \\
\end{boxtable}

\subsection{\Glsfmttext{equipment}}

\subsubsection{Smaller Items}
like arrows or sewing kits may show \pgls{weight} of `$\frac{1}{3}$' or `$\frac{1}{10}$'.
These items take up a regular slot, but you can stack them together; so thirteen arrows have \pgls{weight} of `2', and one arrow has \pgls{weight} of~`1'.

\subsubsection{Tiny Items}
such as `a bit of paper' have \pgls{weight} of `0', but if \pgls{pc} needs to carry over fifty then they should be tracked as having \pgls{weight} of $\frac{1}{100}$.
So four tomes, each with a hundred pages, have \pgls{weight} of~`4'.

\subsubsection{Backpacks}
\index{Backpacks}
let characters hold more items easily, but dropping the \gls{weight} can become tricky in a tight situation.
Dropping a backpack needs a \roll{Dexterity}{Survival} roll at \tn[6] for most backpacks, and the character spends \pgls{ap} as usual.

\subsubsection{In the Arms}
characters can carry a total \gls{weight} equal to half their \glspl{hp} in their arms.
Someone with 8~\glspl{hp} who picks up a bronze statue, with \gls{weight}~12 would receive a -8~Penalty to all actions.
Mounting it on their back reduces the Penalty to~-4.

\subsubsection{To Hand}
means using only one hand, and puts the limit to a quarter of their \glspl{hp}.
If \pgls{pc} with 6~\glspl{hp} lifts a poleaxe with \gls{weight}~4, they would receive a -2~Penalty to attack.
If the character actually hits, the Damage works as usual, so sometimes people find that picking up \pgls{weapon} twice their height feels like the best option they have.

\pic{Roch_Hercka/dwarf_encumbrance}

\subsection{The \Glsfmttext{weight} of Flesh}

\makeAutoRule{bodyweight}{Bodyweight}{triple the maximum \glsfmtplural{hp}}
equals the creature's maximum $\glsfmtplural{hp} \times 3$.
If a human with 7~\glspl{hp} picks up another with 7~\glspl{hp}, they receive a -3~Penalty; the same goes for gnomes and gnolls, or any two people with equal body-sizes, because the \gls{weight} equals their own \glspl{hp} times three.

\index{Caving!Narrow Tunnels}
Narrow cave tunnels often measure who can enter by their \gls{weight}.
The \gls{weight} of items counts towards this total, so characters may not be able to enter a small tunnel until they remove their backpack, or even \gls{armour}.
These items can be pushed through first, or grabbed after.

\subsubsection{Creature Features}
\label{featureWeight}%
\index{Abilities|A}%
such as natural \gls{dr}, wings, a web-spinning thorax, or any other natural ability which add a sizeable feature to an animal's body will have \pgls{weight} to represent how they slow the animal down.
This \gls{weight} does not add to the animal's actual \gls{weight}.

So if you see \pgls{basilisk} \gls{statblock} with a notable \gls{weight}, it does not necessarily indicate items carried, it just shortens the time until the animal becomes exhausted from \glspl{ep}.

\subsection{\Glsfmtlongpl{ep}}
\label{ep}

\Glsentryfullpl{ep} are \pgls{weight} you cannot drop.
They build up slowly as characters exert themselves, threatening quietly, then suddenly present an unfair choice --- should the \gls{pc} drop the last of their food (risking more \glspl{ep} later), stay and rest (in a forest full of curious \glspl{monster}) or endure (and listen to their croaking lungs degrade)?

Everything from combat, to swimming through freezing water, or climbing a cliff, should add \pgls{ep}, but only once the character has stopped, and the adrenaline has worn off.
While mid-combat, characters should only gain \glspl{ep} from spells, not from the fight itself.

\Glsentrylongpl{fp} cannot mitigate \glspl{ep}.
Characters with enough luck to avoid arrows and dragon-fire can still collapse after a long~run.

\subsubsection{Special Categories}
of \glspl{ep} will not disappear with \pgls{interval}'s rest.
Poison and some wicked diseases accrue \glspl{ep} until healed with \pgls{elixir} from the correct \gls{sphere}.
%! Link to making elixirs.

\index{Coins!\Glsfmtlongpl{ep}}
The character sheet holds a special place for a stack of coins to represent \glspl{ep}, below the space for \glspl{hp} coins.
Using different coins to represent different kinds of \gls{weight} and \glspl{ep} helps to track all the changes -- you can feel the weight pinning the character sheet down, as the character's burden increases.
% ...but not on the mini CS.

\end{multicols}

\needspace{18\baselineskip}
\section{Time \& Space}

\begin{multicols}{2}

\noindent
\Gls{travel} is the beating heart and background noise of the game.
\Glsentrydesc{travel}, so standard journeys take a caravan 10 miles per day as they march during the morning and afternoon, then rest throughout the darkness.

A troupe might rush 7~miles each \gls{interval} until night, and cover 21 miles in a day.
But mostly, people don't.
Almost everyone travels together in a caravan, at the pace of their slowest member (the over-burdened horses and donkeys).

Encountering \pgls{monster} happens rarely, despite the fact that \gls{fenestra}'s economic and political structure depend entirely on defending against them and using their bodies.
The \gls{gm} can glide over a journey of 30~miles quickly, by giving a brief description of how the terrain changes, and asking the players to mark off 3~\glspl{ration}.

\subsubsection{\Glsfmtplural{ration}}
\label{hunger}
\index{Hunger}
\index{Starvation}
\index{Food}
abstract every meal and snack into a single token with \pgls{weight} of 1.
The \glspl{pc} rest and chew multiple times a day, but players should only track how many days' of food their character has.

When taverns sell `\rations', or when \glspl{npc} have `\rations', or `\rations' written on the \gls{statblock}, this always counts as `\pgls{ration}', and provides a day's sustenance.

Characters who haven't eaten by night lose \pgls{hp}.
After three days without food, characters will have lost 3~\glsentrylongpl{hp}, and their ability to carry heavy loads on a long march as they become emaciated.

\subsubsection{Dreams}
\index{Sleep Deprivation}
\index{Dreams}
feed the body and soul each night.
Failure to dream for a day inflicts 3~\glspl{ep}.

Troupes sleeping outside at night often take turns to stay awake on \pgls{vigil}.
In this case, the \gls{gm} hands out 3~\glspl{ep} to the players to divide as they wish.
Once character may take 2~\glspl{ep} and remain awake most of the night, or three characters can take \pgls{ep} each, and divide the work evenly.

\needspace{12\baselineskip}
\subsection{Time}
\label{time}

\subsubsection{\Glsfmtplural{round}}
\glsentrydesc{round}.

\subsubsection{\Glsfmtplural{interval}}
\glsentrydesc{interval}.
\index{Doors!Blocking Mana}
\index{Mana!Regeneration}

A central pool of coins makes giving and tracking \glspl{fp} and \glspl{mp} faster than writing and unwriting over the same scuffed square on the character sheet.
\index{Coins!\Glsfmtlongpl{fp}}
The \gls{gm} can simply hand out coins at the end of each \gls{interval}.

\subsubsection{\Glsentrytext{downtime}}
\glsentrydesc{downtime}.
\index{Convalescence}\index{Healing}
Each weekly session covers four weeks in \gls{fenestra}, so the players can declare \pgls{downtime} at any point so characters can rest, research, and heal, as long as the \gls{gm} finds nothing likely to interrupt them.

\label{healing}
When \pgls{pc} has been reduced to 1~\gls{hp}, the player should consider taking a different character for the rest of the session, from their \gls{characterPool} if possible.

The wounded \gls{pc} can go to rest in the nearest \gls{healersGuild}.

\needspace{12\baselineskip}
\subsection{Space}
\label{space}
\index{Space}
\index{Distance}

\subsubsection{\Glsfmtplural{step}}
\index{Steps}
are yards if you like yards, or metres if you like metres.
Or just \glspl{step}.

\subsubsection{Longer Distances}
use these standard approximations, to save the \gls{gm} from having to say `250 metres North'.
In order, they are:

\begin{multicols}{2}
  \begin{enumerate}
    \item
    \toggletrue{Distant}
    \setcounter{spellCost}{2}
    \setRange\stepcounter{spellCost}
    `\spellRange'
    \item
    \setRange\stepcounter{spellCost}
    `\spellRange'
    \item
    \setRange\stepcounter{spellCost}
    `\spellRange'
    \item
    \setRange\stepcounter{spellCost}
    `\spellRange'
    \item
    \setRange\stepcounter{spellCost}
    `\spellRange'
  \end{enumerate}
\end{multicols}

\subsubsection{\Glsfmtplural{area}}
are any place you can see at a glance.
They \glsentrydesc{area}.

\subsubsection{\Glsfmtplural{region}}
\glsentrydesc{region}.%
\exRef{judgement}{Judgement}{threads}

\end{multicols}

\settoggle{genExamples}{false}
