\makeindex[name=spells,title={Spell Summaries},columns=2]

% for CS callout boxes

\newcommand\commentary[3]{
  \node[text=white, overlay, rectangle callout, callout relative pointer={(#2)}, fill=black,] (randomNode) at (#1) [text width=4cm]{#3};
}

\newcommand{\racechart}{

\begin{nametable}[clX]{Race}
  
  \textbf{Roll} & \textbf{Race} & \textbf{Adjustments} \\\hline

  2-3 & Gnome & \mbox{+1 Intelligence, +1 Dexterity,} \mbox{Strength -2, Speed -1} \\

  4-5 & Dwarf & +1 Dexterity, -1 Speed \\

  6-8 & Human & +1 Strength, -1 Wits \\

  9-10 & Elf & +1 Wits, -1 Strength \\

  11-12 & Gnoll & \mbox{+1 Strength, +1 Speed,} \mbox{-1 Intelligence, -2 Charisma} \\

\end{nametable}
}

\newcommand\attributeChart{
  \begin{boxtable}[YY]

    \textbf{Result} & \textbf{Bonus} \\
    \hline

    2 & -3 \\

    3 & -2 \\

    4-5 & -1 \\

    6-8 & 0 \\

    9-10 & +1 \\

    11 & +2 \\

    12 & +3 \\

  \end{boxtable}
}

\newcommand\XPchart{
  \begin{nametable}[XYYYrYY,float*=b]{\Gls{xp} Chart}
    \textbf{Level} &  \textbf{Attributes}  &  \textbf{Skill}  & \textbf{Combat/ \mbox{Projectiles}}  & \textbf{\Gls{fp} Base} &  \textbf{Knacks} & \textbf{Sphere}  \\
    \hline
    <1                &   5        &         &                      &         0 (5 \glspl{fp})  &      & \\
    1st               &   10       &   5     &        10            &        10 (10 \glspl{fp}) &   5  &    10  \\
    2nd               &   20       &   10    &        20            &        15 (15 \glspl{fp}) &  10  &    15  \\
    3rd               &   30       &   15    &        40            &        25 (20 \glspl{fp}) &  15  &    25  \\
    4th               &   50       &         &                      &        45 (25 \glspl{fp}) &  20  &    45  \\
    5th               &            &         &                      &        85 (30 \glspl{fp}) &  25  &    85  \\

  \end{nametable}
}

\newcommand{\initiativechart}{

  \begin{nametable}[Xc]{\Gls{ap} Costs}

  \textbf{Standard Actions} & \textbf{\Gls{ap}} \\
  \hline

  Attacking      & by weapon \\

  Drawing weapon & 1 \\

  Guard Someone & 1 \\

  Ram & 3 \\

  \hline
  \hspace{3em}\textbf{Projectiles} & \\
  \hline

  Crossbow & 1 \\

  Improvised projectile & 2 \\

  Longbow   & 4 \\

  Reloading & 1 \\

  Shortbow & 2 \\

  Thrown weapon & 4 \\
  \hline

  \hline
  \textbf{Other Actions} & \\\hline

  Cast a spell & $1 + Lv - Wts$ \\

  Keeping Edgy & 1 \\

  Moving & 1 \\

  Speaking & 1 \\

  \end{nametable}

}

\newcommand\improvisedWeaponsChart{

  \begin{nametable}[XXXXXX]{Improvised Weapons}

  \textbf{Name} & \textbf{Attack Bonus} & \textbf{Damage Bonus} & \textbf{\Glsfmtshort{ap} Cost} & \textbf{\Gls{weightrating}} & \textbf{Cost} \\\hline

  \boulder & \\

  \chair & \\

  \club & \\

  \cudgel & \\

  \firepoker & \\

  \skillet & \\

  \knife & \\

  \Log & \\

  \rock & \\

  \stick & \\

  \woodaxe & \\

  \end{nametable}


}

\newcommand{\weaponsChart}{
  \begin{nametable}[XXXXXX]{M\^{e}l\'{e}e Weapons}

  \textbf{Name} & \textbf{Attack Bonus} & \textbf{Damage Bonus} & \textbf{\Glsfmtshort{ap} Cost} & \textbf{\Gls{weightrating}} & \textbf{Cost} \\\hline

  \Dagger & 2 \glspl{sp} \\

  \greataxe & 15 \glspl{sp} \\

  \greatsword & 17 \glspl{sp} \\

  \javelin & 5 \glspl{sp} \\

  \longsword & 10 \glspl{sp} \\

  \maul & 8 \glspl{sp} \\

  \poleaxe & 8 \glspl{sp} \\

  \quarterstaff & 2 \glspl{sp} \\

  \shortsword & 8 \glspl{sp} \\

  \spear & 2 \glspl{sp} \\

  \end{nametable}
}

\newcommand\esotericWeaponsChart{
  \begin{nametable}[XXXXXX]{Esoteric Weapons}

  \textbf{Name} & \textbf{Attack Bonus} & \textbf{Damage Bonus} & \textbf{\Glsfmtshort{ap} Cost} & \textbf{\Gls{weightrating}} & \textbf{Cost} \\\hline

  \greatclub & 50 \glspl{cp} \\

  \giantsword & 25 \glspl{sp} \\

  \rapier & 31 \glspl{sp} \\

  \warhammer & 8 \glspl{sp} \\

  \whip & 1 \glspl{sp} \\

  \end{nametable}
}

\newcommand{\shieldchart}{
  \begin{nametable}[XXXXX]{Shields}

  \textbf{Name} & \textbf{Defence Bonus} & \textbf{\Glsfmtshortpl{ap} Cost} & \textbf{Weight} & Cost \\\hline

  \bucklar
  \roundshield
  \kiteshield

  \end{nametable}
}

\newcommand{\armourchart}{

  \begin{boxtable}[LXXXX]

  \textbf{Armour} & \textbf{\Gls{dr}} & \textbf{Weight} & \textbf{Noise} & \textbf{Price} \\\hline

  \textbf{Partial} & & & & \\\hline

  Elvish & 2 & -2 & 0 & 3gp \\

  Padded & 2 & 0 & 0 & 1sp \\

  Leather & 3 & 0 & 0 & 5sp \\

  Chain & 4 &  1 & 1 & 10sp \\

  Plate & 5 &  2 & 4 & 12gp \\
  \hline

  \textbf{Complete} & & & & \\\hline

  Elvish & 2 & -1  & 0 & 9gp \\

  Padded & 2 & 1  & 0 & 3sp \\

  Leather & 3 & 1  & 1 & 15sp \\

  Chain & 4 &  2  & 2 & 30sp \\

  Plate & 5 &  3  & 5 & 36gp \\

  \end{boxtable}
}

\newcommand{\fatiguechart}{

  \begin{nametable}{\gls{fatigue} Chart}
    \textbf{Action} & \textbf{\Glspl{fatigue}} \\\hline

    Armour & Wearing armour inflicts \glspl{fatigue} equal to the armour's \glsentryname{weightrating}. \\

    Bleeding & 1 \gls{fatigue} per slashing damage which was not mitigated by armour. \\

    Climbing & 1 \gls{fatigue} per step. \\

    Fighting & Each round inflicts 1 \gls{fatigue}. \\

    Holding Breath & 1 \gls{fatigue} per round. \\

    Marching & 1 \gls{fatigue} per mile. \index{Marching}\\

    Starving & Each meal skipped inflicts 1 \gls{fatigue} plus half the character's Strength Bonus (rounded up). \\

    Swimming & Each step swum inflicts 1 \gls{fatigue}. \\
  \end{nametable}

}

\newcommand{\moralechart}{
  \begin{nametable}{Morale Chart}
    Bonus & Situation \\\hline

    +4 & Monsters outnumber characters 3:1. \\

    +2 & Monsters outnumber characters 2:1. \\

    +2 & Character's top Strength Bonus is lower than the monster's.  \\

    +2 & Monster is desperate or hungry.  \\

    -2 & Character's top Strength Bonus is higher than the monster's.  \\

    -2 & Characters outnumber the monsters. \\

    -2 & Monster is wounded. \\

    -1 & Players have displayed awesome magical abilities. \\

      \end{nametable}
}

\newcommand{\chasechart}{

  \begin{boxtable}

  Total & Result \\\hline

  11+ & The characters immediately escape their pursuers. \\

  10 & The characters escape their pursuers after travelling through 2 areas. \\

  9 & The characters escape their pursuers after travelling through 3 areas. \\

  8 & The characters are chased through 3 areas and reroll. \\

  7 & The characters are chased through 2 areas and reroll. \\

  6 & The characters are chased through 1 area and reroll. \\

  {\textless}5 & The characters are immediately caught. \\

  \end{boxtable}

}

\newcommand{\huntchart}{

  \begin{boxtable}

  Total & Result \\\hline

  10+ & The characters immediately capture their quarry. \\

  9 & The characters chase their quarry through 3 areas, then capture them. \\

  8 & The characters chase their quarry through 2 areas before catching up with them. \\

  7 & The characters chase their quarry through 1 area and then catch up with them. \\

  6 & The characters chase their quarry through 3 areas, then lose them. \\

  5 & The characters chase their quarry through 2 areas before losing them. \\

  4 & The characters chase their quarry through 1 area, then lose them. \\

  {\textless}3 & The characters immediately lose their quarry. \\

  \end{boxtable}

}

\newcommand{\projectilesChart}{
\begin{boxtable}[XXXXXl]

  \textbf{Projectile} & \textbf{\Gls{ap}} & \textbf{Damage} & \textbf{Weight}  & \textbf{Reload Time} & \textbf{Cost} \\\hline

  Crossbow &  1 & $2D6$ & 1 & 5 - Str \glspl{round}  & {20 sp} \\

  Longbow &  4 & varies & -4 & 4 - extra Str \glspl{ap}& {10 sp} \\

  Shortbow &  1 & $1D6-1$ & -5 & 1 \gls{ap} & {5 sp} \\

  Throwing knives & 2 & $1D6$ & -5 & 1 \gls{ap} & {100 cp} \\

\end{boxtable}
}

\newcommand{\FPRegen}{

  \begin{boxtable}
    Base \glspl{fp} & Regeneration \\\hline

    5 & 2 per scene \\

    10 & 4 per scene \\

    15 & 6 per scene \\

    20 & 8 per scene \\
  \end{boxtable}
}

\newcommand{\animalStats}{
  \begin{nametable}[lccX]{Animal Features}

  \textbf{Animal} & \textbf{Min Str.} & \textbf{Max Str.} & \textbf{Animal Traits} \\\hline

  Badger & -4 & -3 & \quadraped \\

\iftoggle{aif}{
  Basilisk & +5 & +8 & DR 4-6, \quadraped \\
}{}

  Bear & +4 & +5 & DR 2, \quadraped \\

  Beaver & -5 & -4 & \quadraped \\

  Bird/ Bat & -5 & 4 & \flight \\

  Cat & -5 & -5 & \quadraped \\

  Cow & 0 & +4 & \quadraped \\

\iftoggle{aif}{
  Chitincrawler & +3 & +5 & DR 4 \\
}{}
  Deer & 0 & +2 & \quadraped \\

  Donkey & 0 & +3 & \quadraped \\

  Frog & -5 & -5 & \amphibious \\

  Goat & -1 & +2 & DR 2, \quadraped \\

\iftoggle{aif}{
  Griffin & -1 & +2 & \flight \\
}{}

  Horse & +1 & +4 & \quadraped \\

  Large Cat & +1 & +3 & \quadraped \\

  Pig & 0 & +3 & DR 2, \quadraped \\

  Rat & -5 & -5 & \quadraped \\

  Wolf & -2 & +2 & \quadraped \\

  \end{nametable}
}

\newcommand{\servicesChart}{

\begin{tcolorbox}[tabularx={XX},title=Services,bottom=10pt,left=3pt,right=3pt,code={\rowcolors{2}{}{gray!10}}]

  \textbf{Apothacary} & 1 \gls{sp} / level of Intelligence + Medicine \\\hline

  Poison cure & 1 \gls{sp} \\

  Poison & 4 \gls{sp} \\\hline

  \textbf{Fence} & \\\hline

  Common item & 20\% of the item's total value \\

  Identifiable valuables & 40\% of the item's total value \\

  Magical item & Varies considerably, 50\% chance the item is refused \\\hline

  \textbf{Boat Trip} & \\\hline

  Popular trading route & $2D6-2$ \gls{sp} (reroll each day) \\

  Standard stream & $3D6$ \gls{cp} \\\hline

  \textbf{Guide} &  150 \gls{cp}/ day \\\hline

  \iftoggle{aif}{
      Inner villages & 50 \gls{cp} \\
    Near the \gls{edge} & 5 \gls{sp} \\
    Beyond the \gls{edge} & 15 \gls{sp} \\
  }{
    Dangerous area & 1 \gls{sp} \\

    Uncharted area & 50 \gls{cp} \\
  }
  \hline
  \textbf{Hospitality} &  \\\hline

  Public House & 0 \gls{cp} \\

  Meal & 5 \gls{cp} \\

  Barn with breakfast & 10 \gls{cp} \\

  City Inn & 40 \gls{cp} \\

  Fancy Room & 1 \gls{sp} \\\hline

  \textbf{Minstrel} &  15 \gls{cp}/ performance \\\hline

  Large audience & 500 \gls{cp} \\

  Massive audience & 1 \gls{gp} \\

  Creating a new song & 2 \gls{sp} \\

  Illegal song & 5 \gls{sp} \\\hline

  \textbf{Sellsword} & 10 \gls{sp}/ day \\\hline

  Opponent & \gls{xp}$^3$ \gls{sp} \\

  Illegal Murder & 10 \gls{sp} \\\hline

  \textbf{Tracker} &  5 \glspl{sp} / day \\\hline

  \iftoggle{aif}{
    Inner villages & 1 \gls{sp} \\
    Near the \gls{edge} & 10 \gls{sp} \\
    Beyond the \gls{edge} & 20\gls{sp} \\
  }{
    Dangerous area & 2 \gls{sp} \\

    Uncharted area & 4 \gls{sp} \\
  }\hline

\end{tcolorbox}
}

\newcommand\stackingDamageChart{
  \begin{nametable}[l|YYYYYYYYYY]{Damage Conversion}

    \textbf{Damage} & 1     & 2     & 3     & 4   & 5     & 6      & 7     & 8 & 9 & 10 \\
    \textbf{Roll}   & 1D6-3 & 1D6-2 & 1D6-1 & 1D6 & 1D6+1 & 1D6+2  & 1D6+3 & 2D6 & 2D6+1 & 2D6+2 \\

  \end{nametable}
}

\newcommand\tempoChart{
  \begin{nametable}{General Tempo Chart}
    \textbf{Roll} & \textbf{Result} \\\hline
    6 & Encounter, and reroll! \\
    5 & Encounter. \\
    4 & Encounter. \\
    3 & 1 day peace. \\
    2 & 2 days peace. \\
    1 & 3 days peace. \\
  \end{nametable}
}

\renewcommand\csComments{
    \commentary{[xshift=10,yshift=-7em]TCBPOSTER@title.north}{-2em,2em}{{\huge\ref{sumCCrace}}: Write down your name and roll a random race (\autopageref{raceRoll}).}

    \commentary{[xshift=4em,yshift=-6em]TCBPOSTER@attributes.south east}{-4.1em,2.8em}{{\huge\ref{sumCCatt}:} Roll to determine each of your Attributes, then apply the racial modifier (\autopageref{randomAttributes}). The large `\statDot' represents a Bonus of `0'.}

    \commentary{[xshift=13em,yshift=-6em]TCBPOSTER@title.west}{-2em,2em}{{\huge\ref{sumCCconcept}:} Concept and Name: write two words to summarize the character, and a name.}

    \commentary{[xshift=4em,yshift=-1em]TCBPOSTER@gumption.west}{-20em,9em}{}
    \commentary{[xshift=4em,yshift=0em]TCBPOSTER@gumption.west}{1em,-28em}{}
    \commentary{[xshift=4em,yshift=0em]TCBPOSTER@gumption.west}{-4em,6.5em}{{\huge\ref{sumCCxp}:} Spend 50 \glspl{xp} to purchase Attributes, Skills, and Knacks (\autopageref{xp}).}

    \commentary{[xshift=3em,yshift=-1em]TCBPOSTER@gumption.south}{.4em,5em}{\small Place coins in the circles to keep track of the current score.}

    \commentary{[xshift=3em,yshift=2em]TCBPOSTER@spheres.south}{3em,2em}{\small Place a coin on the \glspl{ap} tracker to keep track of your \glspl{ap} during combat.}

    \commentary{[xshift=-7em,yshift=-4em]TCBPOSTER@title.east}{-2em,1em}{{\huge\ref{sumCCcode}:} Select a Code (\autopageref{gods_codes}) and write down a culture.}

    \iftoggle{aif}{}{
      \commentary{[xshift=-8em,yshift=4em]TCBPOSTER@equipment.south}{-7em,-3em}{\small {\huge\ref{sumCCcoin}:} Roll $3D6 - 5$ \gls{cp} starting money, a multiply it by 2 for every Skill-dot you have.}
    }
    \commentary{[xshift=5em,yshift=5em]TCBPOSTER@armoury.north west}{4em,-25em}{}
    \commentary{[xshift=5em,yshift=5em]TCBPOSTER@armoury.north west}{-1em,-2.5em}{{\huge\ref{sumCCequip}:} \iftoggle{aif}{Check your reaons for enlisting, and take your starting equipment (\autopageref{enlistment_equipment}).}{Take a number of pieces of equipment equal to all Skill levels added together (\autopageref{goods}).}}
}


\newcommand\acquisitionXP{

  \begin{xpchart}{Acquisition}

    1 & Buying an expensive item -- worth 1 \gls{gp} or more. \\

    1 & Being greeted deferentially by a stranger. \\

    1 & Confirming you are the richest person in a new region. \\

    1 & Hiring a new servant. \\

    1 & Meeting a famous person. \\

    3 & Buying a very expensive item -- at least 10 \gls{gp} in value. \\

    3 & Disarming a potential conflict. \\

    5 & Starting a new business. \\

    5 & Finding a magical item. \\

    5 & Gaining a new title such as a guild master or townmaster. \\

    5 & Buying an extremely expensive item -- at least 20 \gls{gp} in value. \\

    10 & Marrying into a prestigious family. \\

    10 & Acquiring a horde of new wealth -- at least 100\% of the character's maximum lifetime total, and at least 1 gp. \\

    10 & Buying an expensive new home -- at least 200 \gls{gp} in value. \\

  \end{xpchart}
}

\newcommand\justiceXP{
  \begin{xpchart}{\Glsentrytext{justicegod}}

    1 & Donating at least 1 \gls{gp} to the temple. \\

    1 & Enforcing a law. \\

    1 & Feeding the poor. \\

    1 & Hosting guests. \\

    1 & Punishing law breakers. \\

    1 & Returning someone's valuables to them. \\

    1 & Enduring a troubling oath. \\

    3 & Enforcing a major law or imposing the law on a group. \\

    3 & Donating at least 10 \gls{gp} to the temple. \\

    3 & Returning from a dangerous mission with all allies unharmed. \\

    5 & Completing a quest you swore to complete over one session ago. \\

    5 & Correcting the local law. \\

    5 & Creating a peace treaty between factions in danger of fighting. \\

    10 & Creating a peace treaty between warring factions. \\

    10 & Donating at least 100 \gls{gp} to the temple. \\

    15 & Deposing a tyrant. \\

  \end{xpchart}

}

\newcommand\joyXP{
  \begin{xpchart}{\Glsentrytext{joygod}}

    1 & Playing a prank. \\

    1 & Donating at least 1 \gls{gp} to the church. \\

    1 & Drinking and eating to excess. \\

    1 & Making a particularly bad pun. \\

    1 & Giving food or shelter to the needy. \\

    1 & Reminding another player about something in their Code to get them \glspl{xp}. \\

    3 & Winning a drinking competition. \\

    3 & Lifting the spirits of the downtrodden. \\

    3 & Creating a funny song. Requires at least a full night and an Intelligence + Performance action, \gls{tn} 12. \\

    3 & Playing a prank set up last session. \\

    5 & Hosting a feast for a village. \\

    5 & Creating a new type of alcohol. \\

    5 & Saving someone's life. \\

    10 & Saving a village or larger populated area from destruction. \\

    15 & Playing a prank set more than two sessions ago at the start of the campaign. \\

  \end{xpchart}


}

\newcommand\knowledgeXP{
  \begin{xpchart}{\Glsentrytext{knowledgegod}}

    1 & Donating at least 1 \gls{gp} to the temple. \\

    1 & Learning a new secret. \\

    1 & Correcting someone on world-lore. \\

    1 & Gaining a new level in Academics or any sphere in the Path of Alchemy. \\

    1 & Remembering an \gls{npc}'s name. \\

    1 & Crafting a new magical item. \\

    1 & Drawing a map for the current session. \\

    1 & Overcoming a tricky situation. \\

    3 & Solving a complicated puzzle. \\

    5 & Donating at least 10 \gls{gp} to the temple. \\

    5 & Uncovering a conspiracy. \\

    5 & Writing an informative book on some topic. Intelligence + Academics is rolled at \gls{tn} 16 during downtime, and requires at least a month. \\

    10 & Donating at least 100 \gls{gp} to the temple. \\

    10 & Finding and preserving important knowledge that would otherwise have been destroyed forever. \\

    15 & Solving a legendary puzzle. \\

  \end{xpchart}


}

\newcommand\experienceXP{
  \begin{xpchart}{Experience}

    1 & Finding a new type of food or drink. \\

    1 & Witnessing a flower open. \\

    1 & Seeing a change in Season. \\

    1 & Greeting a new type of humanoid. \\

    1 & Entering a new area, days away from anything seen before. \\

    1 & Hearing a new language. \\

    1 & Learning a new type of creative skill, such as Crafts or Performance. \\

    3 & Experiencing any emotion to heights never reached before. \\

    3 & Learning a Skill or sphere to a level above any other Skill or sphere you have. \\

    3 & Making a new friend. \\

    3 & Returning home after a longer period away than ever before. \\

    3 & Finding a mana lake. \\

    3 & Seeing a new magical item activate. \\

    5 & Experiencing deep tragedy. \\

    5 & Taking \gls{hp} Damage for the first time. \\

    5 & Falling in love. \\

    5 & Creating a masterpiece of some kind at \gls{tn} 14.
    Each Skill can only be used to gain \gls{xp} once. \\

    5 & Discovering a lost piece of art or literature. \\

    10 & Finding an area lost to all contact for more than a century. \\

    15 & Finding an area never before visited by people. \\

  \end{xpchart}
}

\newcommand\natureXP{
  \begin{xpchart}{\Glsentrytext{naturegod}}

    1 & Donating at least 1 \gls{cp} to the temple. \\

    1 & Hunting one's own food and dedicating it to \glsentrytext{naturegod}. \\

    1 & Gaining a new level in the Caving, Seafaring, or Wyldcrafting Skills. \\

    3 & Building a shrine -- requires 3 days work and an Intelligence + Crafts action, \gls{tn} 8. \\

    3 & Donating all of one's money to the temple. \\

    3 & Freeing a wild animal from captivity. \\

    3 & Destroying an `unnatural' creature such as the undead\iftoggle{aif}{, or nura}{}. \\

    5 & Composing a song to \gls{naturegod} -- requires an Intelligence + Performance action, \gls{tn} 15. \\

    5 & Seeing a new type of creature. \\

    10 & Establishing a new temple. \\

    15 & Saving some miles of land from being despoiled. \\

  \end{xpchart}


}

\newcommand\warXP{
  \begin{xpchart}{\Gls{wargod}}

    1 & Donating at least 1 \gls{gp} to the temple. \\

    1 & Finding an interesting battle trophy. \\

    1 & Finishing a battle with 0 \gls{fp}. \\

    1 & Finishing a round with -10 \glspl{ap} or less. \\

    1 & Fighting a new type of creature. \\

    1 & Gaining a new level of the Combat Skill. \\

    1 & Surviving a skirmish while outnumbered. \\

    1 & Going first in the party when entering a dangerous situation. \\

    3 & Answering a one on one challenge. \\

    3 & Killing three opponents single handedly in one battle. \\

    3 & Killing a more dangerous opponent than ever before (danger is measured in \gls{xp}). \\

    3 & Surviving a large scale battle while outnumbered. \\

    5 & Donating at least 10 \gls{gp} to the temple. \\

    5 & Killing five or more opponents single handedly in battle. \\

    5 & Killing a dangerous opponent (16+ \gls{xp}) single handedly. \\

    10 & Killing a dangerous opponent (10+ \gls{xp}) without wearing armour. \\

    10 & Donating at least 100 \gls{gp} to the temple. \\

    10 & Defeating a previously victorious opponent. \\

    15 & Starting a war. \\

  \end{xpchart}


}

\newcommand\deathXP{
  \begin{xpchart}{\Glsentrytext{deathgod}}

    1 & Donating at least 1 \gls{gp} to the temple. \\

    1 & Cursing a dying enemy, to prevent them from achieving a proper afterlife. \\

    1 & Gaining a new level in the Necromancy sphere. \\

    1 & Giving someone blessings upon their death bed. \\

    1 & Losing \glspl{hp}. \\

    1 & Toasting to absent friends. \\

    3 & Performing an outlandish burial, with sacrifices and words appropriate for the deceased. \\

    3 & Easing the last moments of a dying person. \\

    3 & Erecting a shrine to the dead. Requires an Int + Craft action, \gls{tn} 9, at at least 10 gp. \\

    3 & Learning the history of a new area. \\

    3 & Protecting the graves of the honoured dead. \\

    5 & Donating at least 10 \gls{gp} to the temple of \gls{deathgod}. \\

    5 & Falling below 0 \gls{hp}. \\

    5 & Retrieving a dead body for a proper burial. \\

    10 & Building or funding a grand mausoleum. Takes a year and requires at least 100 gp. \\

    10 & Falling below -3 \glspl{hp}. \\

    15 & Falling below -5 \glspl{hp}. \\

  \end{xpchart}

}

\newcommand\tribeXP{
\begin{xpchart}{the Tribe}

  1 & Helping out a member of the tribe. \\

  1 & Greeting a member of the tribe during a long time away. \\

  1 & Welcoming a friend into the tribe. \\

  1 & Testing a potential new member of the tribe. \\

  1 & Defending your tribe's honour. \\

  3 & Spreading the honourable name of the tribe to outsiders. \\

  3 & Entering battle simply for the sake of the tribe. \\

  5 & Forming an alliance for the tribe. \\

  5 & Returning home after an extended trip away. \\

  5 & Acquiring a new home for your family. \\

  5 & Saving a member of the tribe from some terrible situation. \\

  5 & Fulfilling the final wishes of an ancestor.
  The ancestor's wish can be specified only while spending \glspl{storypoint}, and the \gls{xp} can only be collected if the mission is completed in a later Story. \\

  10 & Starting a family. \\

  15 & Saving one's entire tribe from destruction. \\

\end{xpchart}

}

\newcommand\pathsSummary{
  \begin{boxtable}[lp{.25\textwidth}X]

    Path & Spheres & Flavour \\\hline

    Alchemy & Conjuration, Invocation, Force, Illusion, Necromancy & Alchemists use sacred geometry and the power of precious metals and minerals to twist the world around them. \\

    Blood & Aldaron, Enchantment, Force, Invocation, Polymorph & Creatures with innate magic simply call to the world to change the weather, change targets' species, and move items with the power of their minds.  It is used by elves, dragons, and sorcerers with elven blood. \\

    Devotion (\glsentrytext{naturegod}) & Aldaron, Conjuration, Fate, Polymorph & \glsentrytext{naturegod} blesses rare priests of the forest with the ability to change local weather conditions, and cast divine light. \\

    Devotion (\glsentrytext{justicegod}) & Aldaron, Enchantment, Fate, Force & Followers of \gls{justicegod} channel their god to protect the innocent and righteous with blessings and raw magical force.  Evil creatures can be detected, then be ordered to stop, turn and flee. \\

    Runes & Conjuration, Fate, Force, Illusion, Necromancy & Rune magics take a long time as the spells must literally be painted or carved into items. The resulting spells are often placed into items for a quick release, or cast ahead of time. Rune magics are powerful but require craft and preparation. \\

    Song & Aldaron, Enchantment, Fate, Illusion & Song magic must be cast slowly, as the spells are literally songs. The spells have subtle effects but song magic is no less powerful than other spheres. \\

  \end{boxtable}
}
