\section*{The Bodies of Giants}

\epigraph{Intro by Irina}

\begin{multicols}{2}

\noindent
You can call me Ishmail, because I'm gonna tell you about whales.

In Europe's 19th century, the whole world depended on one resource.
Men would travel to the other side of the world, then find and kill a Biblical monster.

\subsubsection{Some Whale Basics}

Whales are broadly divided into two kinds: toothed whales and baleen whales.
Toothed whales have teeth, baleen whales have a flexible filter-feeding system, also named baleen.
Baleen is derived from Latin \textit{balaena}, related to Greek \textit{phalaina}, all of which mean `whale', making `baleen whale' -- `whale whale'.

A notable whale species is sperm whale (like \textit{Moby Dick}, or Pearl from \textit{Sponge Bob}).
Inside its square head, there's a highly sought after liquid called spermaceti.
`Sperm whale' is actually short for `spermaceti whale'.
`Spermaceti' means `whale sperm'.
So `spermaceti whale' means `whale sperm whale'.
`Spermaceti' is a thick, waxy substance that is speculated to aid with bouyancy, however it cannot be confirmed because sperm whales hate us now.
But it definitely isn't sperm.

\subsubsection{Whalers}

Killing a whale is not a fun day.
Everyone who does it has to do it.
Bastards and other surplus -- men who have no apprenticeship, or inheritance -- have one clear path to make their living.
They board a whaling boat, and sail.

Once they have a whale in sight, they propel harpoons using gunpowder weapons, or their arms.
The side of their boat swivels out to provide a netted `shelf' to process the whale.

Once a whale is caught, the whalers must work fast and harvest as much as they can before the whale dies, and rot begins.
Some must walk on top of it, and hope to balance well enough to not fall into the icy-cold waters.
The whalers peel its skin like an orange, then extract and boil the blubber into whale oil.

Some whalers collect toothed whales' teeth and carve them to pass the time (it's either that or homoeroticism).

\subsubsection{Bountiful Corpses}

Whale meat gets harvested for food.
Bones are sometimes ground into fertilizer.
Sperm whales get scalped and have spermaceti extracted from the junk
(`junk' is the actual name of the part of their heads that houses the liquid).
Spermaceti is brought to land, where it's chilled over Winter, and later squeezed into sperm oil.
Solid leftovers were bleached and sold as spermaceti wax.

Baleen is also called `whalebone', despite being closer to nails or hair than bone in structure.
Baleen is harvested from the whale's mouth, and used wherever a thin flexible material can be used.
Most famously, it was used for clothes' construction -- corsets, stays, bustles, crinolines, et c.

Ambergris is fossilized bowel secretions of sperm whales.
It's speculated to be a reaction to bowel irritations.

\end{multicols}

\clearpage

\section*{Shifting the Eye-Line}

\begin{multicols}{2}

\noindent
\Gls{fenestra} does not sit well with our modern instincts.
We all understand that aeroplanes and cows have irrevocably changed the planet for the worse.
But in \gls{fenestra}, people fight against a cruel nature.
They burn the forest while celebrating every toppled tree.
They despise their own world.

And so naturally, the perspective-switch which \gls{fenestra} demands can feel unclean.
But you have to understand, their world feels very different.

When people travelled during this era, they would only fear other humans.
They hated wolves for stealing their food, and in some parts of the planet, a few animals had the muscle and teeth to kill a few humans here and there\ldots at least those not paying attention, but no animal has every posed a real danger to humanity on the same level as diseases, starvation, or warfare.
No city ever disappeared because of large predators eating everyone.
Humanity has wandered the earth, fearlessly, selecting the best plots of land they could to grow all the food they wished, and fence in the meat.

Nobody alive will ever seen an auroch, as we hunted the last of the bovine progenitors to extinction around the 13th century.
They stood almost two metres tall, with horns almost a metre long.
Once we had tamed every forest, we ate the last of them.
Now only our cows remain -- unnatural creatures, created through early genetic manipulation.
Nobody planned this genocide -- we simply hunt too well to let large animals live.

Aurochs might have survived if they had the protection of basilisks and giant arachnids.
These might seem like unlikely farmers, but they only eat the slowest of their prey, keeping their numbers `under control'.
And they do the same for humans.

Disease, starvation and war don't trouble the people in \gls{fenestra} much, but families can swell and shrink in much the same way they did in old Europe.
Women may have half a dozen children, but the population never grows beyond the land which people manage to capture.
Just as Europeans once considered the black death, and other plagues, a standard way to die, any market place in \gls{fenestra} will have updates on what ate whom recently.
This leaves that world very similar to ours on-balance, but very strange in all the details.

\begin{exampletext}
  A young man, who does random jobs to get by, with little pay, must dig a grave tonight.
  The town has no full-time grave-diggers, because most people do not leave a corpse.
\end{exampletext}

\begin{exampletext}
  Two men in a bar debate the usefulness of wearing a helmet.
  The first details the standard attacks from every predator, which mostly involve ambushes, and grabbing.

  \begin{speechtext}
    ``Helmets stop you spotting an attack.
    And they don't help when something grabs you.

    In fact, they make everything worse in every way, defending only against bandits.
    But bandits don't want to eat you, and they're much more timid than real predators, so you can pay them off if you need to.
    \emph{And} the helmet still stops you spotting the ambush.
  \end{speechtext}

  The interlocutor has a few good points on the other side, so the debate continues -- should soldiers wear helmets?
\end{exampletext}

\begin{exampletext}
  The newest recruit to the \gls{templeOfBeasts} arrives, tired, hungry, and pregnant.
  Not long before, most people who became pregnant would receive some dispensation to take time off, or would receive permission to work as \pgls{jotter} instead of fighting at the \gls{edge}.
  However, this lead to an explosion of pregnancies within the \gls{guard}, which lead to a group of the hardest bastards to walk \gls{fenestra}'s dark roads.

  Nobody wants any more of \emph{that lot}, so instead of a safe desk-job, pregnant women now receive a spear or crossbow (depending on supplies).
\end{exampletext}

\begin{exampletext}
  A shepherd walks out with his tiny flock.
  Not many can survive on the small grassland which surround the \gls{village}, but enough to make a daily walk worth the effort.

  Archers stand watch from the \gls{village} walls, waiting for hours as the sheep slowly graze.

  No matter what the traders say, shepherds consider their lives more valuable than any of their sheep, so all shepherds make sure to keep at least one lame sheep within the flock.
  They break a leg or two, and make sure the break never completely heals.
  As a result, the sheep limps slow enough for the shepherd to out-run it, and if anything comes out of the forest, looking for a meal, they can take the lame sheep while the shepherd runs.
\end{exampletext}

\begin{exampletext}
  The young \glspl{doula} make their way to the island -- a paradise where people can walk freely, if they walk in groups of a dozen or so.

  Whenever someone spots a predator, the elders call for a general hunt until they slay the beast.
  They dig up Mouthdiggers, burn chitincrawler eggs, and lay traps for woodspies.
  Basilisks do not swim to the island, and griffins rarely fly this far out.

  The island remains blissfully free, most of the year, leaving the inhabitants to walk freely.

  The elders have forbidden raising children anywhere on the island -- a lesson hard-learned after generations of mishaps.
  Those who begin life on the island remain free forever, but walking free on \gls{fenestra}'s mainland means death.
  As a result, those few surviving, carefree, people who grew up in paradise never leave.
\end{exampletext}

\end{multicols}


