\usepackage[indexonlyfirst]{glossaries}
\setacronymstyle{long-short-desc}
\makeglossaries

\makeglossaries

%%% Core Rules Terms

\newglossaryentry{area}{name={Area},description={The basic unit of large spaces. An area is a space made distinct by its features. In a dungeon, each room might count as an area, while out in the open plains a forest might be composed of the local areas: `the centre with the big, felled tree; the river's fork; the priestess's house and the griffins' nesting site}}

	\newglossaryentry{standingspell}{name={Standing Spell},description={A spell which stays put once cast for as long as the caster wants to maintain it},first={\textit{Standing Spell}}}

	\newglossaryentry{quickaction}{name={Quick Action},description={An action which can skip the normal Initiative order but still costs Initiative.  Quick actions can even be performed when someone has a negative Initiative score},first={\textit{Quick Action}}}

	\newglossaryentry{attribute}{name={Attribute},description={One of the six Traits which form the basis of the character -- Strength, Speed, et c}}

	\newglossaryentry{character}{name={Character},description={Anyone in the game world, though `the characters' typically refers to the PCs}}

	\newacronym[description={The smallest unit of currency},shortplural={cp}]{cp}{cp}{Copper Piece}

	\newacronym[description={One gold piece is worth ten sliver pieces},shortplural={gp}]{gp}{gp}{Gold Piece}

	\newacronym[description={-- the person running the game, playing all characters except the PCs, creating the story and making rulings. Everything rests in the hands of the GM}]{gm}{GM}{Games Master}

	\newglossaryentry{downtime}{name={Downtime},description={This is any long period of time between adventures. It gives characters a chance to complete personal tasks and train in highly technical Skills}}

	\newglossaryentry{evasion}{name={The Evasion Factor},description={A rating of a character's ability to avoid Damage. The Evasion Factor is based on a character's Dexterity plus a bonus from their Weapon. The TN to hit any character is 7 plus their Evasion Factor}}

	\newglossaryentry{strike}{name={The Strike Factor},description={A rating of a character's ability to hit people accurately. The Strike Factor is usually based on a character's Combat Skill}}

	\newacronym[description={A measure of how much luck the character has left, used solely to avoid Damage},shortplural={FP}]{fp}{FP}{Fate Point}

	\newacronym[description={The basic measure of a character's health},shortplural={HP}]{hp}{HP}{Hit Points}

	\newacronym[description={The number players need to roll on the dice to succeed in a task},shortplural={TNs}]{tn}{TN}{Target Number}

	\newacronym[description={The ``battery power'' of a magic user, which allows them to power spells},shortplural={MP}]{mp}{MP}{Mana Point}

	\newglossaryentry{init}{name={Initiative Factor},description={The bonus to the character's Initiative score, usually a combination of Speed + Weapon bonus + Combat Skill bonus}}

	\newglossaryentry{miracleworker}{name={Miracle Worker},text={miracle worker},description={A generic term for any magic user which the author occasionally employs in a futile attempt to seem more high-brow}}

	\newacronym[description={An abstract measurement of how much training characters get.  PCs spend XP to buy Traits},shortplural={XP}]{xp}{XP}{Experience Point}

	\newglossaryentry{natural}{name={Natural Roll},description={A natural roll is a roll where the physical dice land on some number. For example, a `natural 2' is where both dice come up facing 1, as opposed to a player gaining the result `2' from rolling a 3 and getting a -1 penalty. Similarly a `natural 12' is when the dice land on a `12' without modification}}

	\newacronym[description={Non Player Character -- anyone in the world played by the GM rather than a player}]{npc}{NPC}{Non Player Character}

	\newglossaryentry{path}{name={Path},description={Each Path of Magic is a different way to cast spells. Each path has its own available spheres of magic and restrictions}}

	\newacronym[description={-- one of the characters run by the people playing the game}]{pc}{PC}{Player Character}

	\newglossaryentry{round}{name={Round},description={A round is an abstract measurement of time during which characters can make a series of attacks or cast spells. Each new round players adjust their combat tactics}}

	\newglossaryentry{scene}{name={Scene},description={A narrative measurement of time. Each time the players decide to do something new or move somewhere else, it's a new scene. The end of combat always prompts a new scene. While scenes, unlike squares, are variable units of time, the GM is free to set a scene as some definite unit of time, such as `half an hour'}}

	\newacronym[description={Shield Points -- magical shielding from the force sphere},shortplural={SP}]{SP}{SP}{Shield Points}

	\newacronym[description={A rating of protection, generally from wearing armour},shortplural={DR}]{dr}{DR}{Damage Resistance}

	\newacronym[description={One silver piece is worth one hundred copper pieces},shortplural={sp},text={silver piece}]{sp}{sp}{Silver Pieces}

	\newglossaryentry{sphere}{name={Sphere},description={One of the types of magic. Each sphere has 5 levels, each more difficult than the last to obtain}}

	\newglossaryentry{square}{name={Square},description={An abstract unit of measurement. We can imagine it about two yards long, as wide as the squares on your gaming board, or any other length. A more story-based game, without a board, might imagine each square is a `zone' or room -- it matters little so long as each square is a consistent size}}

	\newglossaryentry{trait}{name={Trait},description={Any gaming stat, such as a character's maximum MP, a Skill or an Attribute}}


%%%%% Adventures in Fenestra Side Quests

\newglossaryentry{spiderqueen}{name={The Spider Queen},text={the Spider Queen},description={is an old elf who collects pet chitincrawlers.  She has also started to shapeshift to look like them}}

\newglossaryentry{nightguard}{name={Darren},text={Darren},description={is an active and ambitious member of the night guard},first={Darren of the Night Guard}}

\newglossaryentry{alemaster}{name={Anatole},text={Anatole},description={is the highest ranking member of the temple of Alass\"{e}, who are synonymous with the Ale Guild in this region},first={Anatole, Master of the Ale Guild}}

\newglossaryentry{beardedale}{name={Jafeth},text={Jafeth},description={is the leader of the Ale Guild in the Bearded Mountains, and became rich selling watered-down dwarvish ale},first={Master Jafeth of the Bearded Mountains' Ale Guild}}

\newglossaryentry{krelth}{name={Krelth},description={is a curious gnomish illusionist, who often gets in trouble studying things he shouldn't}}

\newglossaryentry{elfprince}{name={Prince Arinel},text={Arinel},first={Prince Arinel},description={an elf, famed for his songs}}

\newglossaryentry{necromancer}{name={Antonym},first={Antonym, Fallen Priest of Qualme},description={is a priest of Qualme, from the time of Noland Beard.  He has since died, but kept himself alive in order to study Song Magic.  Unfortunately, death has robbed him of his voice}}

\newglossaryentry{forestpriest}{name={Adrian},text={Adrian},description={is a half-elvish high priest of Laique, from the time of Noland Beard.  He has become irritated at how far men are encroaching into the forest, and hopes to stop them}}

\newglossaryentry{baron}{name={Arthur, Townmaster},text={Lord Arthur},first={Jacob Arthur (local Townmaster)},description={is an evil landowner, currently attempting to become an Areamaster by building outwards into the local forest.  He has already found the location of a once-lost city, and plans on rebuilding it}}

\newglossaryentry{cleverbaron}{name={Catelina},text={Lord Catelina},first={Villagemaster Catelina},description={is a villagemaster, currently plotting to overthrow the king by collecting magical items from the nura}}

\newglossaryentry{banditking}{name={Redclaw},first={Redclaw Whiteland},description={was the child of a noble in Whiteland, but his family were killed by the king.  He's lived as a criminal ringleader ever since}}

\newglossaryentry{sewerking}{name={Areth},first={Areth Whiteland},description={is a Whiteland noble, currently living under the town}}

\newglossaryentry{sewerthief}{name={David},first={David Whiteland},description={misses the comforts of nobility.  He spends a lot of his time running errands through \glsentrytext{pig} and across town, such as delivering cursed magical items to \glsentrytext{banditking}}}

\newglossaryentry{nurabaron}{name={Clandon, the Villagemaster},text={Villagemaster Clandon},first={Clandon, Villagemaster to Redfall,},description={was a jovial and simple Villagemaster until he had an accident with a magical item, and has since turned into an ogre}}

\newglossaryentry{captain}{name={Captain Oscar Quenn},text={Captain Oscar},first={Captain Oscar of the Night Guard},description={is the captain of the Night Guard in town}}

\newglossaryentry{patron}{name={The Patron},text={the Patron},first={the Patron underground},description={a mysterious man who came through the portal under the town, from the Realm of Darkness and Fire.  He has been making deals about magical items, in return for food}}

\newglossaryentry{king}{name={King Wyatt Fenestra},text={King Wyatt},first={King Wyatt Fenestra},description={keeps a tight control over Fenestra through the use of magical portals}}

\newglossaryentry{pigowner}{name={Wendy, Owner of \glsentrytext{pig}},text={Wendy Pig},first={Wendy, owner of \glsentrytext{pig}},description={a middle-aged, fierce woman who runs the toughest bar in town}}

\newglossaryentry{bard}{name={Roger},text={Roger},first={Roger the bard},description={is a young and headstrong man who writes better than he sings}}

\newglossaryentry{townmaster}{name={Townmaster Marcus},text={Lord Marcus},first={Marcus the townmaster},description={rules the local town, and takes steep taxes from the local Villagemasters.  He's an approachable and fun fellow, but generally hated by everyone who knows his tax collectors better than they know him}}

%%%%% Adventures in Fenestra Places

\newglossaryentry{redfall}{name={Redfall},first={Redfall (a village in the North)},description={is rather out-of-the-way village, famed for its red-earth clay and waterfall}}

\newglossaryentry{college}{name={The Collage of Alchemy},text={the Collage of Alchemy},first={the Collage of Alchemy},description={is where all alchemists go for training.  They maintain a complete monopoly on alchemical magic, and all spells must be learned through them or one of their outposts}}

\newglossaryentry{pig}{name={The Mincing Pig},text={the Mincing Pig},first={the Mincing Pig Tavern},description={boasts the reputation as the loudest tavern in town, placed close to the entrance.  Many people don't get farther than `the Pig' when coming to town}}

\newglossaryentry{whitehorse}{name={The White Horse},text={the White Horse},first={the White Horse Inn},description={is a serene and esteemed establishment, with a reputation for high-quality clientel}}

\newglossaryentry{lostcity}{name={The Lost City},text={the Lost City},first={the Lost City},description={is a city where men lived centuries ago.  It is now so destroyed and overgrown that people do not know they have entered until they see some part of an old stone building}}

\newglossaryentry{shatteredcastle}{name={The Shattered Castle},text={the Shattered Castle},description={is where the king lives.  It has six sections in six different regions, and each one contains a magical gateway to the hub in Whiteland, so different parts of the  castle exist in different places at once}}


%%%%% Gnolls Revolt

\newglossaryentry{thshen}{name={Thshe\'{n}},description={A gnoll slave who became educated to see if gnolls could learn.  He intends to free every gnoll he can from slavery}}

\newglossaryentry{donald}{name={Townmaster Donald},description={A townmaster with the kind of confidence only complete ignorance can bring}}

\newglossaryentry{darren}{name={Darren of the Gnolls' Guild},description={Darren is the Journeyman for the Gnolls' Guild, visiting the area to sell the services of more gnolls}}

