\chapter[Chamber of Magic]{Magic}
\label{spellsChapter}
\toggletrue{examplecharacter}
\index{Spells}

\begin{multicols}{2}

\noindent
Magic can do any number of things, but like anything else, it has standards.
These standards are `spells'.

\Glspl{casting}
\glsentrydesc{casting}.

\Glspl{spell}
\glsentrydesc{spell}.

\subsubsection{The Elemental \glsfmtplural{sphere}}
are Combat \glspl{skill}, just like Projectiles and Melee.
To cast a spell, check the `Requirements'; if it says `Air 2', then the character needs the `Air' \gls{skill} at level 2 to learn and cast it.

\subsubsection{The High \glsfmtplural{sphere}}
are a mix of the lower.
Fire and Air make `Light', Fire and Earth make `Force'.
Characters don't have to learn anything extra to learn these High \glspl{sphere} -- they only need the two Elemental \glspl{skill}.

High \glspl{sphere} show their requirements as their base elements, so when you look up a `Death~2' spell, it will show the requirements as `Air~2, Fate~2'.
As long as you have both of those at level 2 or above, you can cast the spell.

\end{multicols}

\section{Elemental \Glsfmtplural{sphere}}

\begin{multicols}{2}

\subsection{Air}
\label{airSpells}

% Reorder spells so they can be alphabetical

\setcounter{diceNo}{0}%

\showSpells{Air1}

\showSpells{Air1}

\spell{Sleep Fog}% Name
  {Devious}% Enhancements
  {Wane}% Action
  {Air}% Spheres
  {ventilation}% Resist with
  {With a memory of a bad night's sleep, the caster thinks about suffocating air.
  Within one \showOnset, the air in the room turns rancid, inflicting \arabic{spellPlusTwo}~\glspl{ep} over the course of a few hours}% Description
  {Anyone inside rolls \roll{Strength}{Vigilance} (\tn[10]) to wake up each time they take another \gls{ep}.}

\showSpells{Air2}

\showSpells{Air2}

\spell{Clawing Wisps}% Name
  {Detailed, Duplicated}% Enhancements
  {Wax}% Action
  {Air}% Spheres
  {wind speed}% Resist with
  {The caster gives a battle-cry, and cloud-lined spirits, appearing like small copies of the caster, push any enemies (but not allies) back, reducing their \glspl{ap} by \arabic{spellPlusOne}}% Description
  {}

\showSpells{Air3}
\showSpells{Air3}

\showSpells{Air4}

\subsection{Fire}
\label{fireSpells}

\showSpells{Fire1}

\showSpells{Fire2}

\spell{Toothless Fire}% Name
  {Duplicated}% Enhancements
  {Warp}% Action
  {Fire}% Spheres
  {intensity of the flame}% Resist with
  {This spell changes the basic nature of a fire, so it can no longer burn a certain type of material}% Description
  {Casters generally use this to stop a fire damaging people, but it could just as easily stop a fire spreading to clothing.

    The fire will not stop feeding on whatever allows it to currently burn, so a wood fire will continue burning its wood.}

\showSpells{Fire3}% Brilliant Bull


\spell{Traitorous Camp}% Name
  {Devious, Distant}% Enhancements
  {Warp}% Action
  {Fire}% Spheres
  {\roll{Wits}{Survival}}% Resist with
  {The caster gives a jealous stare at a fire at \spellRange.
  After two \glsfmtplural{interval}, the fire gains some weight, and starts to burn low, and look strange.
  After a full day, the fire collapses on top of the nearest person, inflicting \showDam~Damage}% Description
  {}

\spell{Naked Fire}% Name
  {Duplicated, Distant}% Enhancements
  {Warp}% Action
  {Fire}% Spheres
  {size of fire}% Resist with
  {After instructing all the fire around on proper etiquette, fires with a total area up to \arabic{spellTargets} \glspl{step} at \spellRange\ turn blue, and stop burning flesh.
  They continue to act upon wood, hair, and clothing as normal}% Description
  {Large fires are more difficult to affect, as they're too busy eating to listen.}

\spell{Direct Lightning}% Name
  {Detailed, Distant, Duplicated}% Enhancements
  {Warp}% Action
  {Fire}% Spheres
  {\roll{Wits}{Survival}}% Resist with
  {The caster directs an active lightning storm to strike \arabic{spellTargets} people, inflicting \showDam\ Damage}% Description
  {The spell targets clouds above, at \spellRange, but those clouds can strike targets at any distance nearby.}

\showSpells{Fire4}

\iftoggle{verbose}{
  \subsection{Earth}
  \label{earthSpells}

  \showSpells{Earth1}

  \showSpells{Earth1}

  \showSpells{Earth2}

  \showSpells{Earth2}

  \spell{Mud Clamp}% Name
    {Duplicated, Detailed}% Enhancements
    {Wax}% Action
    {Earth}% Spheres
    {\roll{Strength}{Survival}}% Resist with
    {The caster shouts `\textit{hold fast}', as snow or mud becomes solid, encasing any feet pushed into the ground.
    Targets then count as \textit{Entangled}}% Description
    {
    The spell suffers a -2~Penalty when used against creatures without flat feet (such as hooves, or most other animal feet), but can also gain up to a +2~Bonus in particularly deep mud or snow.}

  \showSpells{Earth3}

  \showSpells{Earth4}
}{}

\subsection{Water}
\label{waterSpells}


\showSpells{Water1}

\showSpells{Water1}

\showSpells{Water2}

\spell{Detect Water}% Name
  {Distant}% Enhancements
  {Witness}% Action
  {Water}% Spheres
  {smallness}% Resist with
  {The caster detects liquid at \spellRange}% Description
  {They cannot tell the type of liquid -- it could be a raging river, a tankard of ale, or a magma stream}

\showSpells{Water2}

\showSpells{Water3}

\showSpells{Water3}

\showSpells{Water4}

\subsection{Fate}
\label{fateSpells}

\showSpells{Fate1}
\showSpells{Fate1}

\showSpells{Fate2}
\showSpells{Fate2}

\spell{Blessed Army}% Name
  {Duplicated, Distant}% Enhancements
  {Wax}% Action
  {Fate}% Spheres
  {targets' Charisma (inverted)}% Resist with
  {The caster gives good wishes to a troupe, and \arabic{spellTargets} targets receive \showDam~\glspl{fp}}% Description
  {The spell targets those with the highest Charisma before others.}

\showSpells{Fate3}

\spell{Wheel of Fate}% Name
  {Devious, Duplicated}% Enhancements
  {Warp}% Action
  {Fate}% Spheres
  {\roll{Charisma}{Survival}}% Resist with
  {The caster wonders what the future holds, and the nearest \arabic{spellTargets} targets become locked into the same encounter, on repeat.
  It begins half way through the \showOnset, then repeats again, until it has gone through \arabic{spellPlusTwo} iterations}% Description
  {\par
  Of course, the encounter cannot contain the same creatures or people each time, but whatever the \gls{gm} rolled for the encounter will repeat.
  This might result in three \glspl{griffin} looking at troupe, then three \glspl{griffin} attacking, then three \glspl{griffin} nesting, and so on.
  Each of these encounters will come in addition to the standard encounters.}

\showSpells{Fate4}

\subsection{Divergent}
\label{divergentSpells}

\spell{Barrel Blast}% Name
  {Divergent}% Enhancements
  {Wane}% Action
  {Air, Water}% Spheres
  {\roll{Speed}{Cultivation}}% Resist with
  {As the caster spits, \arabic{spellPlusTwo} barrels of water (or equivalent) turn to vapour and rush around a target.
  If the target breathes in, they gain \arabic{spellPlusTwo}~\glspl{ep}}% Description
  {
    The water should be clearly visible, such as a stream or an actual barrel of water.}

\spell{Dock Hickory}% Name
  {Divergent}% Enhancements
  {Warp}% Action
  {Fire, Water}% Spheres
  {\roll{Speed}{Athletics}}% Resist with
  {As the caster sings a song in reverse, nearby fire falls from its wood and flows across the floor like water, while water gathers in place and swirls like flame}% Description
  {
  The fire only deals \showDam, spread evenly across the nearest four \glspl{step}; it dissipates at the end of the \gls{round}, unless it has fuel spread about the ground.
  The water-flame remains upright until the wind eats through its peripheral `sparks'.
  }

\spell{Dark Doom}% Name
  {Divergent}% Enhancements
  {Wane}% Action
  {Fate, Fire}% Spheres
  {\roll{Charisma}{Survival}}% Resist with
  {As the caster bids farewell to someone, they lose \showDam~\glspl{fp}, while the fire closest to the target dies}% Description
  {
  The spell can destroy any fire up to those large enough to deal \showDam\ Damage, and larger fires grow dim.}


\iftoggle{verbose}{
  \spell{Miner's Bane}% Name
    {Divergent}% Enhancements
    {Wane}% Action
    {Earth, Air}% Spheres
    {\roll{Speed}{Cultivation}}% Resist with
    {As the caster curses the target's bread, air, and Sunlight, the ground underneath their feet softens, and their next breath of air sticks to the inside of their throat like thorny cotton.
    Targets who fail to dart away quickly fall prone, and take \arabic{spellPlusTwo}~\glspl{ep} unless they held their breath}% Description
    {
    This spell relies on soft ground to make the target slip.
    Harder floors will not melt, and the spell will settle on the nearest soft earth it can find.}

  \spell{Finding Courage}% Name
    {Divergent}% Enhancements
    {Wax}% Action
    {Earth, Fate}% Spheres
    {malleability of the earth}% Resist with
    {Reaching into some soft ground, such as snow or mud, the caster begs for a chance to fight, and the ground forms into a weapon.
    The caster then gains \arabic{spellPlusTwo}~\glspl{fp}.
    The shatters if it deals over \arabic{spellPlusTwo} Damage}% Description
    {
    The weapon can be anything the caster has used before, up to \pgls{weight} of \arabic{spellPlusTwo}}

}{}

\iftoggle{verbose}{

  \spell{Asking Rock}% Name
    {Detailed, Divergent}% Enhancements
    {Warp}% Action
    {Earth, Fate}% Spheres
    {civilization rating}% Resist with
    {The caster sells a weapon to a rock in return for summoning a beast, and the rock becomes magnetic.
    It has an effective Strength of \arabic{spellPlusTwo} to hold any iron weapon placed next to it.

    Once the rock has a weapon, the \gls{gm} makes an encounter roll}% Description
    {
      The \gls{tn} depends on how close to civilization the spell is cast.
      Attracting \glspl{monster} to \pgls{village} makes more of a challenge than attracting them while past the \gls{edge}.%
      \exRef{judgement}{Judgement}{civilizationRating}
      If the encounter roll produces nothing, the \gls{gm} goes up the chart until they find an encounter.

      \Glspl{npc} targeted by this spell will survive if only their \gls{cr} value is greater than the \gls{monster} which finds them.
      Otherwise, they end in the jaws of a beast.
    }

}{}

\spell{River Smog}% Name
  {Divergent, Duplicated}% Enhancements
  {Wane}% Action
  {Air, Water}% Spheres
  {\roll{Speed}{Athletics}}% Resist with
  {The caster spits prophecies of surreal catastrophe, and \arabic{spellTargets} worth of water turn into a putrid vapour.
  Anyone who breathes it in receives \arabic{spellPlusTwo}~\glspl{ep}}% Description
  {
    The water should be clearly visible, such as a stream or an actual barrel of water.}

\spell{Night Terrors}% Name
  {Distant, Divergent}% Enhancements
  {Wane}% Action
  {Fate, Fire}% Spheres
  {\roll{Charisma}{Survival}}% Resist with
  {The caster calls the first line of a dirge from the darkness, then a fire at \spellRange\ goes out.
  Whoever held it loses \showDam~\glspl{fp}}% Description
  {}

\showSpells{FireWater3}

\showSpells{FireFate4}

\showSpells{WaterAir4}

\end{multicols}

\section{High \Glsfmtplural{sphere}}

\begin{multicols}{2}

\subsection{Light}
\label{lightSpells}

\spell{Little Big}% Name
  {}% Enhancements
  {Warp}% Action
  {Fire, Air}% Spheres
  {\roll{Wits}{Vigilance}}% Resist with
  {The caster mixes a malaphor, and the target's apparent size contorts like a fun-house mirror, inflicting a -\arabic{spellPlusOne} Penalty to ranged attacks against them}% Description
  {A gnome seen from ten steps might appear to be the size of a human, or a human may appear the size of a gnome.
    The closer one approaches the target, the more normal it appears.}

\showSpells{Light1}

\showSpells{Light1}

\spell{Shadow}% Name
  {}% Enhancements
  {Wane}% Action
  {Fire, Air}% Spheres
  {\roll{Wits}{Vigilance}}% Resist with
  {As the caster whispers, a single patch of darkness, the length of a step, forms anywhere within range}% Description
  {It can obscure a candle's light, and diminish a torch's light to the strength of a candle.

    The patch of shadow joins to a target.
    If cast on a person, the shadow remains attached to them, inflicting a -\arabic{spellPlusOne}~Penalty to noticing things (and becomes quickly irritating).

    The shadow fades once sunlight touches it for \pgls{interval}.}


\showSpells{Light2}

\pic{Roch_Hercka/flashing_light}

\showSpells{Light2}

\spell{Shadow Puppet}% Name
  {Detailed}% Enhancements
  {Wane}% Action
  {Fire, Air}% Spheres
  {\roll{Wits}{Vigilance}}% Resist with
  {One target shadow forms a detailed silhouette, and may even stand up and away from the ground}% Description
  {During the daylight, it simply looks weird; but at night these shadow puppets can easily fool someone into thinking \pgls{monster}, or even a particular person lies waiting, in the distance.

  The shadows disappear once bright light hits them.}

\spell{Quiet Campfire}% Name
  {Devious}% Enhancements
  {Wane}% Action
  {Fire, Air}% Spheres
  {size of fire}% Resist with
  {The caster stares at a fire, and thinks dark thoughts.
  After a couple of hours, the fire's light dims, though the noise and heat continue.
  The spell lasts one \showOnset, so the light returns during the night}% Description
  {Travellers sometimes use this spell to hide their fire when camping beyond the \gls{edge}, as it removes at last one chance of an encounter from the fire's light}


\showSpells{Light3}

\spell{Eternal Night}% Name
  {Devious, Duplicated}% Enhancements
  {Wane}% Action
  {Fire, Air}% Spheres
  {\roll{Intelligence}{Survival}}% Resist with
  {The caster pulls on a memory of seeing darkness at the bottom of a clear loch, and the darkness comes.
  After two \glsfmtplural{interval}, a patch of darkness forms, 5~\glsfmtplural{step} across.
  Within four \glsfmtplural{interval}, all vision suffers a -4 Penalty}% Description
  {
  \Glspl{witch} commonly use this spell to make anyone in \pgls{bothy} or other sleeping-space miss the day.
  They awaken late, in darkness, and generally assume they simply cannot sleep.
  After all, they can see darkness, so it must be night, not day.

  By the time people start to question their sanity, and light a fire, the next night may have already come.
  }

\pic{Roch_Hercka/polymorph}

\showSpells{Light4}

\iftoggle{verbose}{

  \subsection{Force}
  \label{forceSpells}

  \showSpells{Force1}

  \showSpells{Force1}

  \spell{Friendly Bolt}% Name
    {Duplicated}% Enhancements
    {Wane}% Action
    {Earth, Fire}% Spheres
    {\gls{weight} carried}% Resist with
    {The caster shouts `let's go', and two targets gain a +\arabic{spellPlusOne}~Bonus to running}% Description
    {They feel suddenly lighter, as if being carried.
    However, this spell does not help with carrying heavy items.}

  \spell{Migrating Butterfly}% Name
    {Detailed}% Enhancements
    {Wane}% Action
    {Earth, Fire}% Spheres
    {\gls{weight} carried}% Resist with
    {The caster outright denies how heavy something is, despite obvious heft, and its \gls{weight} reduces by \arabic{spellPlusOne} when facing towards the direction the caster looks}% Description
    {Items which fall below \gls{weight}~0 will float like a seed on the wind}

  \spell{Traveller's Curse}% Name
    {Distant}% Enhancements
    {Wax}% Action
    {Earth, Fire}% Spheres
    {\roll{Speed}{Athletics}}% Resist with
    {The caster wishes the target a bad journey, and they immediately feel heavier.
    For the next \arabic{spellPlusOne}~\glspl{interval}, the target counts as carrying something with \pgls{weight} of \arabic{spellPlusOne}}% Description
    {
    The curse wears off once the target has marched enough, but any \glspl{interval} the target spends stationary do not count towards the spell's duration.}

  \showSpells{Force2}

  \showSpells{Force3}

  \showSpells{Force4}

  \subsection{Life}
  \label{lifeSpells}

  \showSpells{Life1}

  \spell{Forest Crown}% Name
    {}% Enhancements
    {Warp}% Action
    {Earth, Water}% Spheres
    {\roll{Strength}{Athletics}}% Resist with
    {The target hears how good they would look with antlers, and antlers start to grow.
    They grant +1 to Damage while using the Brawl Skill}% Description
    {
    The antlers will stop the target wearing a helmet, so they cannot wear Complete armour.
    The antlers shed over any cold season, or can be removed with a bone-knife.}

  \spell{Slashers}% Name
    {}% Enhancements
    {Warp}% Action
    {Earth, Water}% Spheres
    {\roll{Strength}{Athletics}}% Resist with
    {The caster makes accusations of vile language, and the target grows needle-sharp claws.
    They grant +1 Damage when using Brawl, but inflict a -1~Penalty to using weapons or other tools}% Description
    {
    The claws grow out naturally, or can be cut off at any time.}

  \spell{Tweak}% Name
    {}% Enhancements
    {Warp}% Action
    {Earth, Water}% Spheres
    {target's \glspl{hp}}% Resist with
    {The caster lists how the target looks like a different race, and their features begin to shift to match}% Description
    {The target must have Attributes within the racial maximums and minimums of the target race, so a human with Strength +1 could become an elf, but Strength +3 would stop them looking remotely elf-like.

    The spell lasts until the target makes use of any characteristic of their own race.
    For humans, this means their ability to walk long-distances, and for gnomes it means trying to focus on anything.}

  \spell{Regeneration}% Name
    {Detailed}% Enhancements
    {Wax}% Action
    {Earth, Water}% Spheres
    {current lost \glspl{hp} + 7}% Resist with
    {The caster gathers tiny gifts for the target, and they begin to regenerate Damage faster, recuperating an extra \arabic{spellPlusOne} \glspl{hp} each week}% Description
    {During this time, the target's appetite grows substantially, demanding an extra \arabic{spellCost} meals every day to keep the spell going.}

  \showSpells{Life2}

  \spell{Scent in the Air}% Name
    {Distant}% Enhancements
    {Witness}% Action
    {Earth, Water}% Spheres
    {target's strangeness}% Resist with
    {Focussing on the air, and sniffing the wind, the caster asks politely if it will say whether or not some creature sits at \spellRange\ in a particular direction}% Description
    {Other humanoids are \tn[7] to detect, beasts are \tn[9], and stranger creatures present a greater challenge.}

  \showSpells{Life3}

  \spell{Herd Protection}% Name
    {Duplicated, Distant}% Enhancements
    {Warp}% Action
    {Earth, Water}% Spheres
    {current \glspl{hp}}% Resist with
    {The caster yells warnings, and the targets gain \gls{dr}~\arabic{spellPlusOne} as their skin becomes thick}% Description
    {Unfortunately, the tough hide feels heavy, and effectively counts as an item with \pgls{weight} of \arabic{spellPlusOne}.
    The hide usually comes with fur, and protects against the cold as well as Damage.

    When the spell combines with others, to boost the target's Strength and grant claws, they eventually look just like a regular bear.

    The strange skin fades to normal once the target takes Damage and heals.}

  \spell{Piglet Farm}% Name
    {Duplicated, Detailed}% Enhancements
    {Wane}% Action
    {Earth, Water}% Spheres
    {\glspl{hp}}% Resist with
    {The caster cries `\textit{Pigs}', and the first \arabic{spellTargets} targets within \spellRange\ shrink, taking a -\arabic{spellPlusOne} Strength Penalty}% Description
    {
    This Penalty stacks as usual, so targets who already have a basic Strength Penalty of -2 would take this Penalty, and then half of their regular Strength Penalty, for a total of -\arabic{spellPlusTwo}.}

  \showSpells{Life4}
}{}

\subsection{Mind}
\label{mindSpells}

\showSpells{Mind1}

\showSpells{Mind1}

\spell{Shadow Check}% Name
  {}% Enhancements
  {Witness}% Action
  {Fate, Water}% Spheres
  {quiet thoughts}% Resist with
  {This spell tells the caster if any particular patch (and it has to be a specific patch) contains anything which has a mind}% Description
  {Shadows, statues, or suspicious bushes can all hide all manner of things.

    The spell will not say what kind of mind this is -- whether animal or person.
    It reports only `yes/ no', and often has false positives due to mice or insects (1/6 chance in daylight, 2/6 chance at night).}


\spell{Awaken}% Name
  {Detailed}% Enhancements
  {Wax}% Action
  {Water, Fate}% Spheres
  {spell's level}% Resist with
  {The caster explains a plan to a spell, and the spell gains an Intelligence Bonus of +\arabic{spellPlusOne} and begins enacting the plan}% Description
  {
  All spells have a small spark of sentience, endowed by their creator, so this spell exploits that fact to make a spell behave more intelligently.

  The spell begins with all of the caster's knowledge, although anything beyond general knowledge generally ends up as a tangled mess of half-memories.
  Without any interest in socializing, any social memories will simply not make sense.
  The spell has all of the caster's Skills which do not depend on any social intuition%
  \footnote{The spell gains none of the caster's magic Spheres},
  with a -1~Penalty.
  However, if given a Charisma Bonus, with the Social Spell (\vpageref{Social Spell}), the spell will have full access to the caster's Skills (again, with a -1~Penalty to each of them).

  The spell grants no Charisma Bonus, and the resulting mind has no interest in anyone, or anything, except as instruments for \emph{the plan}.
  It operates methodically, and sometimes behaves in strange, and alien ways, given its single-pointed value system.%
  \footnote{Look up `the paperclip maximizer' for an explanation of how this kind of spell can easily go horribly wrong.}
  }

\spell{Fear}% Name
  {Detailed}% Enhancements
  {Wane}% Action
  {Water, Fate}% Spheres
  {\roll{Wits}{Academics}}% Resist with
  {The caster explains to the target their imminent death.
    If the target ever finds themselves facing the danger the caster described, they take a -\arabic{spellPlusOne}~Penalty to Mind Attributes}% Description
  {The spell lasts until the character gains \pgls{xp} for following their Code.}

\stepcounter{diceNo}
\showSpells{Mind2}

\pic{Roch_Hercka/elvish_enchanter}

\showSpells{Mind2}

\spell{Social Spell}% Name
  {Detailed}% Enhancements
  {Warp}% Action
  {Water, Fate}% Spheres
  {spell's level}% Resist with
  {The caster speaks to a spell like a dog, which awakens an instinct to socialize, and grants it a +\arabic{spellPlusOne} Charisma Bonus (but no Intelligence)}% Description
  {
  Casters might use this spell to let an illusion interact semi-realistically with people (it will smile and listen, but cannot speak).

  The newly created mind gains all of the caster's Skills which rely primarily on social graces (Deceit, Empathy, and Performance), with a -1~Penalty to each.
  It also gains their spell Spheres with a -1~Penalty, but does not need an audible voice to use them.

  The spell typically tries to attack the caster's enemies, and then interact with the caster and their friends.

  }

\spell{Wishfinder}% Name
  {Detailed}% Enhancements
  {Witness}% Action
  {Water, Fate}% Spheres
  {\roll{Wits}{Academics}}% Resist with
  {The \glsentrytext{witch} commands the target to state their `truest desires', while locking eyes.
    The spell confirms whether or not the target's statement conforms to their Code}% Description
  {The desire must be something which would \textit{directly} grant at least one \glsentrytext{xp}.%
  \exRef{stories}{Stories}{codes}}

\label{carrierCrow}
\spell{Carrier Crow}% Name
  {Detailed, Distant}% Enhancements
  {Warp}% Action
  {Water, Fate}% Spheres
  {\roll{Wits}{Brawl}}% Resist with
  {Whistling to a distant animal, the caster instils an instinct to enact a single plan.
  The animal gains Intelligence +\arabic{spellPlusOne}, used only to complete the plan}% Description
  {
  The spell is most often used to have birds deliver short messages, written on small slivers of paper, but could imbue enough intelligence for any number of missions.

  The target animal gains all of the caster's Skills which do not rely on understanding social graces (so it cannot use any magical Spheres).

  Less cohesive animals, such as a bee-hive or swarm of stirges%
  \footnote{As Terry Pratchett once said, a bee is not an animal -- the hive is the animal, it simply has more mobile cells than most.}
  add +4 to the \gls{tn}.
  }

\showSpells{Mind3}

\showSpells{Mind3}

\showSpells{Mind4}

\subsection{Death}
\label{deathSpells}
\label{Death1}

\showSpells{Death1}
\showSpells{Death1}

\showSpells{Death2}
\showSpells{Death2}

\spell{Corpse Stench}% Name
  {Distant}% Enhancements
  {Witness}% Action
  {Fate, Air}% Spheres
  {time since death}% Resist with
  {Sniffing at the wind, the caster identifies any corpses at \spellRange}% Description
  {Most use this trick to double-check if an immobile and wounded person might still have a chance of life.}

\spell{Infectious Migraine}% Name
  {Duplicated}% Enhancements
  {Wax}% Action
  {Fate, Air}% Spheres
  {\roll{Strength}{Athletics}}% Resist with
  {As the caster roars, a deathly feeling flows over the targets, who gain \arabic{spellPlusOne} \glspl{ep}}% Description
  {}

\spell{Preservation}% Name
  {Detailed}% Enhancements
  {Wane}% Action
  {Fate, Air}% Spheres
  {existing rot}% Resist with
  {The caster paints a picture -- with words or brush -- of food, a shirt, or a human corpse.
  The target ceases all rotting when not in use.
  Rain or snow may fall upon it, but it will not degrade}% Description
  {
    Without this spell, any undead creatures would rot ever faster than a corpse in the ground.

    The spell degrades when the target begins moving, and fails within \arabic{spellPlusOne} \glspl{interval}.
    Therefore, undead spellcasters, such as liches, must continuously cast this spell upon themselves.

    \index{Necromancy!Preservation}
    If cast immediately, upon death, it can arrest the effects of \textit{rigor mortis} before they begin.
  }

\showSpells{Death3}

\spell{Stench Sort}% Name
  {Distant, Detailed}% Enhancements
  {Witness}% Action
  {Fate, Air}% Spheres
  {time since death}% Resist with
  {Sniffing at the wind, the caster identifies any corpses from a particular species at \spellRange\ in the chosen direction}% Description
  {Possible types include `elves', `\glspl{woodspy}', or `bats'.
  The spell only confirms if such a corpse exists.}

\showSpells{Death4}

\subsection{Divergent}

\showSpells{MindLight2}

\showSpells{MindLight3}

\iftoggle{verbose}{
  \showSpells{MindForce4}
}{}

\end{multicols}

