\begin{exampletext}

The next morning the trio gave a wide berth to the area between the fallen village and the mountain, in the hopes of avoiding attack from the rear.
Unfortunately there was little they could do to hide themselves, and a band of hobgoblins from the village were following them.
The trio had reached halfway up the mountain by this point but the hobgoblins were faster than them, and stronger.
Sean thought for a moment about abandoning Hugi if there was a problem - his little dwarvish legs were no good for sprinting.
Of course Hugi's death would not buy them much time, and Arneson would never stand for it.
No, they would have to stick together to survive.

The trio climbed for a while longer, looking back every few moments to note how close the hobgoblins were behind them. An ogre was among their ranks, so they must have come from a very deep cavern.

Looking back, the enemy was nearing and everyone was out of breath.
Arneson suggested a rest to make sure they would be ready for the fight - they could fight downhill against an enemy fatigued from walking upwards.
He calmly got out the rations - some cheese, smoked pork, oatcakes and a flagon of wine.

``May as well have the best of the rations now, eh, friends?'', Arneson said while smiling, and they slowly masticated their age-hardened meal and tried to smile back as the nine foot monstrosity which was so recently a man made its way up to them, pounding its great feet up the mountainous slopes, surrounded by half a dozen hobgoblins, each the size of a broad-shouldered man.

As the hobgoblins neared the plateau where the trio sat they began to make their war cries, but Arneson just sat and ate his last oatcake slowly. They began to sprint upwards across the rocky ground.

The \gls{gm} decides that since the players have the higher ground, they will receive +1 to all rolls until the hobgoblins can reach slightly higher ground - probably after the first action.
One of the players must roll Initiative, so Arneson's player take the dice and rolls, producing a `5'.
Each character's Initiative Bonus adds to this separately, so Sean's Initiative of +1 gives him a total of 6; Hugi's Initiative of +0 gives him a total of 5 and Arneson's Initiative Bonus of +2 makes his total 7.
The group then gain +1 each due to the higher ground.
Meanwhile, the hobgoblins have rolled a `5' for their Initiative as well.
The hobgoblins each have a +1 bonus to Initiative and the ogre has a +0 bonus.

The \gls{gm} knows the highest Initiative total is somewhere under 9 so she calls out,

``Ten! The hobgoblins gather at the base. Nine! They cover their faces with their weapons and gang together. Eight! They push forwards and ....''

``Eight!'' shouts Arneson's player. ``I'm going at eight! \ I'm going for the ogre - you said he was unarmoured so we should be able to take him down with a couple of good hits''.

Arneson's player needs to roll an 8 to hit the ogre, so he rolls and adds +1 for his Combat Skill. His total is `11' and he hits. He has already rolled his damage die at the same time. It landed on a `2'. He adds +2 for his Strength and +1 for the sword's Damage bonus for a total of 5. The ogre shrieks in pain as Arneson's sword sticks in. Arneson's action took 6 Initiative points so he goes down to `2', and the hobgoblin has spent 2 Initiative to defend himself.

``Seven!'', she cries.
Sean's player jumps in.
He only deals $1D6+2$ Damage but he has a much better Combat Skill of +2.
He rolls to Strike but misses the ogre with a `5'.
His Initiative score reduces to 1, and he reduces the hobgoblin's from 6 to 4.

``Six!'', beckons the \gls{gm}, and starts to describe leering hobgoblins stabbing at everyone's feet from the base of the great stone step the trio are sitting on. She gives them each a -1 to attack due to occupying the lower ground.

\pic{Boris_Pecikozic/nura_brawl}{\label{boris:brawl}}

A hobgoblin hits Arneson, so he spends 2 Initiative to attempt to Dodge. Arneson rolls a 7 but the \gls{tn} is 10.
The hobgoblin's Strength is +2 and the battle axe adds +3 Damage for a total of +5.
4 of the Damage is replaced with a die, so the hobgoblin is rolling $2D6+1$ Damage.
The total is 6.
Arneson's player first reduces that by his chainmail's \gls{dr} of 4, leaving 2 Damage.
Instead of taking that Damage he marks off 2 \glsentrylongpl{fp} and declares that the attack in fact misses.

Hugi finally releases his crossbow, but in all the confusion misfires. He's down to Initiative 1.

\needspace{3cm}
\begin{wrapfigure}{r}{.3\textwidth}

	\begin{tabularx}{.3\textwidth}{c|X}
		\setcounter{enc}{12}

		8 & Arneson deals 5 Damage to a hobgoblin. \\

		7 & Sean misses. \\

		6 & Most hobgoblins attack. \\

		5 & Ogre grabs Hugi, then runs away. \\
		4 & Two hobgoblins attack. \\

	\end{tabularx}

\end{wrapfigure}

``Four!'', shouts the \gls{gm}, then she smiles. The next hobgoblin attacks and Arneson rolls a 5 -- that's a failure with a Margin of 4, so it bypasses his chainmail. The axe is coming down towards his unarmoured shin-bone and the Damage rolled is 9. He marks off his last 9 \gls{fp} rather than taking any Damage. Any further Damage is coming straight off his \gls{hp}.

``The ogre pushes forward with its club then reaches out to grab Hugi. Roll at \glsentrylong{tn} 9''.

Hugi's player isn't happy, as that is enough to hit him, or in this case grab him. In fact with his crossbow out rather than a defensive weapon, he'll have a hard time defending himself.

The ogre reached forward, grabbing Hugi by the beard and pulling him back through the horde of hobgoblins and out from the protection of his companions.

The other players want to attack the ogre, but he's making a movement action -- these count as \gls{quickaction} so they are allowed to operate before other actions.
Hugi disappears behind the crowd.
The ogre's Initiative reduces to 0 and he is done for the \gls{round}, but Arneson and Sean each have one action left.

\end{exampletext}

