\begin{exampletext}

	Arneson decides he is going to \textit{ram} the crowd of hobgoblins to save his friend.
	Pushing the half dozen hobgoblins back is going to be tricky.
	He launches himself from the stony step they are on, pushes his chest into one then grabs two more hobgoblins.
	Since he is pushing back 2 extra figures, he takes a -2 penalty to the action.
	Arneson's rolling with +2 from his Strength Bonus and +1 bonus for his Combat Skill for a grand total of +1.
	The \gls{tn} is 7 plus the hobgoblins' Strength of +2 and Combat Skill of +2 for a total of 11.
	The \gls{gm} allows him a total of a +2 bonus for jumping off the step.
	The dice come up with an 8 and his total of +3 just passes the test.
	Normally, he would only push the hobgoblins back by 1 step, but they are on the side of a cliff and being pushed onto their back feet.

	The \gls{gm} decides some sort of check is in order to see how well the hobgoblins perform.
	Ordinarily, she would roll for each of them but there are six of them and that will take too long.
	Thinking quickly - because who wants to slow down combat? - she decides that all of them could potentially fall down the cliff since the first three are in front of the next three so Arneson is pushing against all of them one way or another.
	She gives them a \gls{tn} of 9 to stay up and a bonus of +6 because there are 6 of them.
	Each Margin they roll in the final score is one hobgoblin that has not fallen over.
	Dice clatter, she has rolled a `4' and that leaves a final score of 10.
	Everyone falls down the mountain's steep incline except for a single hobgoblin.

	Sean, on Initiative 1, is the last to act.
	He jumps off the cliff-side to attack the last hobgoblin.
	He strikes with a score of 11, bypassing the ugly creature's Partial chain armour, then rolls $1D6+2$ for the Damage for a total of 4.
	The creature is reduced to half its \glspl{hp} with a crimson gash across its throat.

	As Sean's sword swooped down it opened up his target's arm. The last one standing cries out and withdraws his arm then backs off.

	``End of the \gls{round}!'', cries the \gls{gm}. ``Round two! Roll for Initiative''.
\end{exampletext}

