\begin{exampletext}


The players start by assigning their Combat Skill.
On the last \gls{round}, they left it as the default - it added to the Strike Factor.
Arneson repeats the move and Hugi has no Combat score to speak of, but Sean has Combat +2.
He knows speed is of the essence if he wants to save his friend, so he adds +2 to his Initiative Factor, giving him a total of +3.

The characters roll to get their bearings but achieve only a `4', so Sean will act at Initiative 7.
The \gls{gm} rolls for the hobgoblins and achieves `9' - with Speed +1 they will act on Initiative 10.

\vspace{1em}

\begin{tabularx}{.3\textwidth}{c|X}
\setcounter{enc}{12}

10 & The Ogre grapples Hugi. \\

7 & Sean moves to stab the Ogre. \\

6 & The Ogre deals 5 Damage. \\

5 & Sean kills the ogre. \\

\end{tabularx}

\vspace{1em}

``Twelve!'', the \gls{gm} rolls a Morale Check for that last hobgoblin. It is wounded and outnumbered. The \gls{tn} is 12 and it can add its Combat bonus, but the roll still fails.

``The last hobgoblin backs up. \textit{Eleven!} It flees down the mountain towards its allies, many of whom are still rolling down the hill.''

``Ten!'', the \gls{gm} continues, and immediately rolls for the ogre as it tries to eat Hugi's face off.
This will count as a grappling roll, so he and Hugi will use double their Strength Bonus added to their Combat Skill.
Unfortunately Hugi has neither Strength Bonus nor Combat Skill, so the ogre gets a straight +12 bonus; the roll succeeds before it is even made, and succeeds by a margin of 3: that means Damage is inflicted, rather than just Fatigue Points.
The ogre only adds Strength - of course his massive club is useless for the attack.
His Strength of +5 means he will roll $2D6+1$ Damage for a total of 4. Hugi is safe for now as the ogre luckily bites down on dwarvish helmet as Hugi's player marks off 4 \glspl{fp}.

``Nine! The ogre pulls Hugi down. Eight! He bites down on Hugi's face but gets a mouthful of helmet instead. Seven!''
Sean's player is acting now and takes two Initiative to run over to aid Hugi.
He asks the \gls{gm} if he can sneak up on the ogre.

``You mean in the middle of a fight you want to backstab someone?''

``Sure. He's busy eating Hugi's face, so can I stab him while he's not bothering to avoid it?''

The \gls{gm} thinks about it - the action is not clearly covered in the rules, so she decides the following.

``Okay - make a sneak roll.
If he sees you then he's going to stop the action and defend himself, otherwise your next attack can count as a Sneak Attack. Roll Speed + Stealth at \glsentrytext{tn} 6.''

Sean has no bonus to either, but that ogre is so dim the test is easily passed. 

``Six! The ogre gnaws into Hugi's face, this time without failure. 5 Damage!''

Hugi's player marks off his last 3 \glspl{fp} then 2 \glspl{hp}, noting that he could have just died.

Arneson runs over to aid the fight.

``Five! Sean rolls for attack''

His Sneak Attack gives him +4 to strike the ogre - which he does - and +2 Damage, making his Damage roll $2D6$. His total is 7 Damage.

``Launching himself forward he lands the tip of his sword into the ogre's back just as teeth are sinking into the dwarf's face.
It finds purchase and slides in only six inches before stopping.
The giant whirls \gls{round}, ripping the sword out and pushing Sean to the side.
He screams and attempts to get up, then slumps back down onto the dwarf, blood pooling out of the gash on his back.''

``Finally!'', shouts Sean.
We're done.
It's finished.
We can\ldots

``Four! Over a dozen hobgoblins can be seen marching down from the mountain.'', shouts the \gls{gm}.

``What? \ We can't handle any more. Hugi's Damaged. Arneson's in poor shape too.''

``Three! \ They pull out crossbows and start cranking them {\dots}''

\end{exampletext}
