\begin{exampletext}

Sean's next plan was to run away as fast as possible.
This was much more popular than his last plan of yelling at the enemy, though Hugi found it hard to move enthusiastically with an arrow in his shoulder.
The hobgoblins of course realised that they had no time to reload so they just gave chase.
They dropped their projectiles, pulled out their shortswords and started to clamber along the rocky face of the mountain.
The trio could not move clearly up the mountain until they had gained some ground between themselves and the hobgoblins, and feared that there would be more openings to once-dwarven tunnels, now infested with hobgoblins, if they went further up from their present location.

The basic \gls{tn} for such actions is 6 and the \gls{gm} lowers it by 2 because the trio have a good head start.
The hobgoblins add their Speed Bonus of +1 for a final \gls{tn} of 5.
The party roll an 8 but unfortunately Hugi isn't the fastest of people -- he's only four feet tall after all -- so his score is 7.
They needed that 8 to completely get away.
Arneson and Sean decide they're going to keep pace with him rather than running ahead.
They're not caught yet, but run through three different distinct areas before making another roll.

The hobgoblins were fast on their trail as they clambered over the rocky mountain side.
They soon headed up steeper, overhanging rocks and at one point had to help each other upwards across large rocks jutting out of the side as the hobgoblin horde came ever closer.
Finally, they reached the peaks and gazed down the other side.
Seeing only mist in the other side they decided to lose themselves in the crevices there before the enemy could catch up enough to see their direction.

Arneson's player wants to roll again while adding his +2 Stealth Skill. He is the only one with this Skill but the \gls{gm} says he can use it to help everyone hide.  Arneson's player takes the dice and rolls; he scores 11.

They didn't go far, but only hopped down a few stony crevices before Arneson beckoned them to the side and requested they creep into a nook he had found.
Hugi and Sean could only just fit, with no spare room, so Arneson then bounded off to see what else he could find.
He was still out looking for a spot when the great axes scraping down the cliff could be heard, and guttural voices complained about dangerously empty stomachs.

\end{exampletext}
