\begin{exampletext}
``That's the end of the scene'', the \gls{gm} says. ``You can each regain 2 \glsentrylongpl{fp}.''

``I've got 10 \glsentrylongpl{fp} in total'', mentions Arneson's player, ``So I'm getting 4. But doesn't this rest period count as a new scene too?''.

``Sure, says the \gls{gm}. ``Mark down another load for hiding in the tops of the mountains.''

With their \gls{fp} now replenishing quickly, the group can rest and worry less about being hit again.

``Oh! I've been forgetting about the Fatigue'', says the \gls{gm}. Your \gls{gm} will probably say the same at some point.

``Everyone got two Fatigue from being in two \glspl{round} of combat, then four more for running through three areas.
That's six in total''.

\vspace{1em}

\needspace{3em}

\textbf{Hit Points and Fatigue}

\begin{tabularx}{\linewidth}{ccccccccccc}

$\CIRCLE$ & $\CIRCLE$ & $\CIRCLE$ & $\CIRCLE$ & $\CIRCLE$ & $\CIRCLE$ & $\CIRCLE$ & \Circle & \Circle & \Circle \\ 

$\boxslash$ & $\boxslash$ & $\boxslash$ & $\boxslash$ & $\boxslash$ & $\boxempty$ & $\boxempty$ & $\boxempty$ & $\boxempty$ & $\boxempty$ \\ 

&&&&&&& -1 & -2 & -3 \\

\end{tabularx}

\vspace{1em}

Sean's player starts by putting down two \glspl{fatigue} for the two rounds of combat.
His armour has a \gls{weightrating} of 0, so it doesn't add \gls{fatigue}.
He receives three \gls{fatigue} for running across the mountain.
Five Fatigue Boxes are marked down in total.
If the characters had continued being active that would be the end, but since they have finished the scene while resting, Sean heals 4 Fatigue points leaving him with only 7 in total.

Hugi isn't doing so well. He only had 2 \glspl{hp} left by the time he was running.
He gains the full 7 \glspl{fatigue}.
Finally, the \gls{gm} reminds him that he is bleeding from his wounds.
He is in no condition to patch them up while hiding, especially since nobody in the party knows anything about Medicine.
She decides to only award one more Fatigue point since the arrow is also stopping the wound from bleeding too much - that makes the total 8.
Hugi's rest allows him to regenerate 3 Fatigue Points (he's not as strong as Sean) so he receives 5 Fatigue Points in total.
Dwarves, luckily, can withstand 2 additional Fatigue Points so 2 of those points give him no penalty.
That's 6 more than his \glspl{hp}.
He gains a -4 penalty to all actions.
The danger now passed, the warriors lie in their hiding nooks, watching the cold clouds whirl around them, hoping to never see any kind of goblin again.
They breathe in and out gently, waiting for the heaviness in the chest to subside.
Despite the winds, Sean can hear a gentle drip, drip, drip from the slowly bleeding wound on Hugi's shoulder where an arrow still lies.

\end{exampletext}

