\begin{exampletext}

	Sean wants to know what exactly that spell-song the boy was talking about means.
	He has the Academics Skill and never wrote down any specialisations so he asks the \gls{gm} if he can take a specialisation in whatever it is he needs in order to know about that bard and what he was up to.
	The \gls{gm} tells him to write down `Music' -- it's a fitting specialisation since he already plays the flute.

	Sean then needs to roll using his Intelligence Bonus and Academics Skill -- the first is at 0 but his Academics Skill grants him +1.
	The more academics you have, the better the chances that one of them knows something, so Hugi's player wants to help using his +2 Bonus.
	He has Intelligence +1 and the Academics Skill at first level.
	He wants to use his specialisation in History, and the \gls{gm} allows it.
	Hugi is helping the action so he can add half his score of +3 (rounded up) for an additional +2 bonus.
	Arneson has no knowledge of academic matters and an Intelligence Attribute at -1 so he's better off staying quiet.
	Adding the bonuses together, the roll is at +3.
	The \gls{gm} stipulates the \gls{tn} of 14 -- a tall order without a library to aid matters.
	The roll fails and the troop will have to go into the situation blind.

	Sean recalled what the old alchemist had said about the book -- surely the bard had sung the song from the stolen book and put the guards to sleep with this magic.
	He searched his mind for where such a song might come from and what else it might be capable of.
	He wondered if the bard had really come from the South and what he might want with that book.
	The entire thing was an impenetrable enigma.

\end{exampletext}

